---
title: "glmms_land_cover_change(y/n)_and_occurrences"
author: "Beatrice Trascau"
date: "2024-05-31"
output: html_document
---

# **3.1. GLMMS: Land Cover Change (Y/N) and Occurrence Records**

This markdown outlines the process of modelling the number of species occurrence records in relation to land cover changes, time and SSB ID.

```{r library load, include = FALSE}
library(here)
library(dplyr)
library(lme4)
library(DHARMa)
library(glmmTMB)
library(pscl)
library(ggplot2)
```

```{r data load, include = FALSE, eval = FALSE}
load(here("data", "occurrences_SSB_land_cover.rda"))
occ_SSB_land_cover <- occurrence_SSB_df
```

```{r prepare df for analyis, include = FALSE, eval = FALSE}

# 1. First period of change: 2000-2006 -------------------------------------------------

# Prepare data for the period 2006-2009
occ_df_after_2000.2006 <- occ_SSB_land_cover |>
  select(V1, gbifID, year, species, land_cover2000, land_cover2006, 
         SSBid, cell_ID, TARGET_FID) |>
  filter(!is.na(land_cover2000) & !is.na(land_cover2006)) |>
  filter(year >= 2006 & year <= 2009) |>
  mutate(cover_change = if_else(land_cover2000 == land_cover2006, "N", "Y"))

# Calculate number of records for the period 2006-2009
occ_df_after_2000.2006_records <- occ_df_after_2000.2006 |>
  group_by(cell_ID) |>
  summarise(
    ocurrences_after = n(),
    land_cover2000 = first(land_cover2000),
    land_cover2006 = first(land_cover2006),
    cover_change = first(cover_change),
    SSBid = first(SSBid)) |>
  mutate(time_period = "2000-2006") |>
  select(-c(land_cover2000, land_cover2006))

# 2. Second period of change: 2006 - 2012 ----------------------------------------------

# Prepare data for the period 2012-2015
occ_df_after_2006.2012 <- occ_SSB_land_cover |>
  select(V1, gbifID, year, species, land_cover2006, land_cover2012, 
         SSBid, cell_ID, TARGET_FID) |>
  filter(!is.na(land_cover2006) & !is.na(land_cover2012)) |>
  filter(year >= 2012 & year <= 2015) |>
  mutate(cover_change = if_else(land_cover2006 == land_cover2012, "N", "Y"))

# Calculate number of records for the period 1997-2000
occ_df_after_2006.2012_records <- occ_df_after_2006.2012 |>
  group_by(cell_ID) |>
  summarise(
    ocurrences_after = n(),
    land_cover2006 = first(land_cover2006),
    land_cover2012 = first(land_cover2012),
    cover_change = first(cover_change),
    SSBid = first(SSBid)) |>
  mutate(time_period = "2006-2012") |>
  select(-c(land_cover2006, land_cover2012))

# 3. Second period of change: 2012 - 2018 ----------------------------------------------

# Prepare data for the period 2015-2018
occ_df_after_2012.2018 <- occ_SSB_land_cover |>
  select(V1, gbifID, year, species, land_cover2012, land_cover2018, 
         SSBid, cell_ID, TARGET_FID) |>
  filter(!is.na(land_cover2012) & !is.na(land_cover2018)) |>
  filter(year >= 2015 & year <= 2018) |>
  mutate(cover_change = if_else(land_cover2012 == land_cover2018, "N", "Y"))

# Calculate number of records for the period 1997-2000
occ_df_after_2012.2018_records <- occ_df_after_2012.2018 |>
  group_by(cell_ID) |>
  summarise(
    ocurrences_after = n(),
    land_cover2012 = first(land_cover2012),
    land_cover2018 = first(land_cover2018),
    cover_change = first(cover_change),
    SSBid = first(SSBid)) |>
  mutate(time_period = "2012-2018") |>
  select(-c(land_cover2012, land_cover2018))

# 4. Combine the above 3 into a single df ----------------------------------------------

occ_df_after_records <- bind_rows(occ_df_after_2000.2006_records,
                                  occ_df_after_2006.2012_records,
                                  occ_df_after_2012.2018_records)

```

## Attempt 1: Model 1: Number of occurrences after change by land cover, time period and SSB ID - negative binomial

```{r model 1, include = FALSE, eval = FALSE}

# Run negative binomial model
model1_nb <- glmer.nb(ocurrences_after ~ cover_change * time_period + (1|SSBid),                    data = occ_df_after_records)

# Save model output to file to save time next time
save(model1_nb, file = here::here("data", "models", "model1_nb.RData"))
```

### Model Validation

#### Convergence Check

```{r convergence check}

# Load model from file
load(here::here("data", "models", "model1_nb.RData"))

# Get summary of the model
summary(model1_nb)
```

Model output suggests a significant effect of time and a significant interaction between time period and land cover change (Y) but no significant effect of land cover. Some high correlation values between cover_change_Y and cover_change_Y:2006_2012 as well as cover_change_Y and cover_change_Y:2012_2018. But this is likely to be expected as cover_change_Y is nested into cover_change_Y:2006_2012 and cover_change_Y:2012_2018. Most residuals (from scaled residuals) are close to zero. But the maximum residual (= 534.64) is very large. The AIC and BIC values are also very large.

#### Simulate Residuals with DHARMa

```{r model residuals}

# Simulate residuals
model1_nb_simulation <- simulateResiduals(fittedModel = model1_nb)

# Plot simulated residuals
plot(model1_nb_simulation)
```

Simulated residuals deviate significantly from expectations, are overdispersed and have many outliers. The within-group deviations from uniformity are significant. Test for homogeneity of variance is significant.

#### Residuals Diagnostics

```{r residuals diagnostics}

# Perform residual diagnostics
testResiduals(model1_nb_simulation)

# Plot residuals
#plot(residuals(model1_nb))
```

Everything that could go be with theresiduals has gone wrong.

#### Check Random Effects

```{r random effects}

# Check distribution
#ranef(model1_nb)
```

#### Test for Overdispersion

```{r test for overdispersion}
testDispersion(model1_nb_simulation)
```

Residuals are significantly overdispersed.

#### Test for Zero-Inflation

```{r test for zero-inflation}
testZeroInflation(model1_nb_simulation)
```

Significant zero inflation. Attempting a zero-inflated negative binomial model to address these issues.

## Attempt 2: Model 1: Number of occurrences after change by land cover (Y/N) and SSB ID - zero inflated negative binomial

```{r model 2. include = FALSE, eval = FALSE}

# Run zero inflated negative binomial model
model2_nb_zero <- glmmTMB(ocurrences_after ~ cover_change * time_period + (1|SSBid),
                          ziformula = ~ 1,
                          family = nbinom2, 
                          data = occ_df_after_records)

# Save model output to file to save time next time
save(model2_nb_zero, file = here::here("data", "models", "model2_nb_zero.RData"))
```

### Model Validation

#### Convergence Check

```{r convergence check}

# Load model from file
load(here::here("data", "models", "model2_nb_zero.RData"))

# Get summary of the model
summary(model2_nb_zero)
```

Model output suggests a significant effect of time and a significant interaction between time period and land cover change (Y) but no significant effect of land cover. AIC and BIC values still large. Dispersion parameter = 0.50.

#### Simulate Residuals with DHARMa

```{r residuals, include = FALSE}

# Simulate residuals 
model2_nb_zero_simulation <- simulateResiduals(fittedModel = model2_nb_zero)

# Plot simulated residuals
plot(model2_nb_zero_simulation)
```

Simulated residuals deviate significantly from expectations, are overdispersed and have many outliers. The within-group deviations from uniformity are significant. Test for homogeneity of variance is significant.

#### Test for Overdispersion

```{r test for overdipserison}
testDispersion(model2_nb_zero_simulation)
```

Test for overdispersion is significant.

#### Test for Zero-Inflation

```{r test for zero-inflation}
testZeroInflation(model2_nb_zero_simulation)
```

Test for zero inflation is significant. Clearly, this model is not usable either.

## Attempt 3: Model 1: Number of occurrences after change by land cover change (Y/N) and SSB ID - Non-zero inflated negative binomial with glmmTMB

```{r model 3. include = FALSE, eval = FALSE}

# Run negative binomial model
model3_nb <- glmmTMB(ocurrences_after ~ cover_change * time_period + (1|SSBid),
                          family = nbinom2, 
                          data = occ_df_after_records)

# Save model output to file to save time next time
save(model3_nb, file = here::here("data", "models", "model3_nb.RData"))
```

### Model Validation

#### Convergence Check

```{r convergence check}

# Load model from file
load(here::here("data", "models", "model3_nb.RData"))

# Get summary of the model
summary(model3_nb)
```

Model output suggests a significant effect of time and a significant interaction between time period and land cover change (Y) but no significant effect of land cover. AIC and BIC values still large. Dispersion parameter = 0.50.

#### Simulate Residuals with DHARMa

```{r simulate residuals}

# Simulate residuals 
model3_nb_simulation <- simulateResiduals(fittedModel = model3_nb)

# Plot simulated residuals
plot(model3_nb_simulation)
```

Simulated residuals deviate significantly from expectations, are overdispersed and have many outliers. The within-group deviations from uniformity are significant. Test for homogeneity of variance is significant.

#### Test for Overdispersion

```{r test for overdispersion}
testDispersion(model3_nb_simulation)
```

#### Test for Zero-Inflation

```{r test for zero inflation}
testZeroInflation(model3_nb_simulation)
```
