---
title: "glmms_land_cover_change(y/n)_and_occurrences"
author: "Beatrice Trascau"
date: "2024-05-31"
output: html_document
---

# **3.1. GLMMS: Land Cover Change (Y/N) and Occurrence Records (Simple Model)**

This markdown outlines the process of modelling the number of species occurrence records in relation to land cover changes, time and SSB ID.

The models ran as part of this are as follows:

| Model Number | Function | Family  | Link  | SSBID/Municipality | Ran? |
|--------------|----------|---------|-------|--------------------|------|
| 1.1          | glmer.nb |         |       | SSBID              |      |
| 1.2          | glmmTMB  | nbinom1 | log   | SSBID              |      |
| 1.3          | glmmTMB  | nbinom1 | log   | Municipality       |      |
| 1.4          | glmmTMB  | nbinom1 | logit | SSBID              |      |
| 1.5          | glmmTMB  | nbinom2 | logit | Municipality       |      |

```{r library load, include = FALSE}
library(here)
library(dplyr)
library(lattice)
library(lme4)
library(DHARMa)
library(glmmTMB)
library(mgcv)
library(gamm4)
library(ggplot2)
```

## Data Exploration

### Inspect Data

```{r inspect data}

# Read in data
load(here::here("data", "derived_data",
                "occ_y_n_cover_change_after_records_for_model.rda"))

glimpse(occ_y_n_cover_change_after_records_for_model)
```

### Check for Outliers

```{r check outliers}

# Barplot of number of occurrences
ggplot(occ_y_n_cover_change_after_records_for_model , 
       aes(x = cover_change, y = ocurrences_after, fill = time_period))+
  geom_boxplot()+
  theme_classic()
```

There are some very clear outliers in the number of occurrence records, more so for the pixels where the land cover doesn't change.

### Check Normality

```{r check normality}

# Plot histogram of number of records by time period
occ_y_n_cover_change_after_records_for_model |>
  filter(ocurrences_after < 100) |>
  ggplot(aes(x = ocurrences_after)) +
  geom_histogram(binwidth = 2, position = "dodge", 
                 alpha = 0.4) +
  facet_grid(time_period ~ ., scales = "free") +
  theme_classic()
```

### Check Zero Inflation

```{r check zero inflation}

# Calculate % of rows with values of 0 for number of occurrences
100 * sum(occ_y_n_cover_change_after_records_for_model$ocurrences_after == 0) / nrow(occ_y_n_cover_change_after_records_for_model)
#suspicious
```

### Check for Collinearity

Covariates in model (cover change, SSB ID and time period) are all categorial. No need to check for collinearity between them.

### Check Relationships

```{r check relationships}

# Plot relationships
occ_y_n_cover_change_after_records_for_model |>
  ggplot(aes(x = time_period, y = ocurrences_after)) +
  geom_point() +
  facet_wrap(~cover_change) +
  theme_classic()


```

Relationships difficult to determine but seems that in the case of the pixels where there is land cover change, the relationship with time is not linear.

## Model 1.1.

The model: ocurrences_after \~ cover_change \* time_period + (1\|SSBID) failed to converge: "Warning: Model failed to converge with max\|grad\| = 0.00724284 (tol = 0.002, component 1)Warning: Model is nearly unidentifiable: very large eigenvalue - Rescale variables?"

The model ocurrences_after \~ cover_change \* time_period + (1\|municipality) also failed to converge: "Warning: Model failed to converge with max\|grad\| = 0.00214456 (tol = 0.002, component 1)Warning: Model is nearly unidentifiable: very large eigenvalue - Rescale variables?"

This model ran with no problems when the finer SSB ID grid was previously used.

## Model 1.2.

### Model Validation

#### Convergence Check

```{r convergence check}

# Load model from file
load(here::here("data", "models", "model1.2_nb_SSB.RData"))

# Get summary of the model
summary(model1.2_nb_SSB)
```

Model output suggests a significant effect of time and a significant interaction between time period and land cover change (Y) but no significant effect of land cover. AIC and BIC values still large. Dispersion parameter = 45.7.

#### Check Residuals

```{r model 1 attempt 2 check residuals}

# Create tibble for plotting
plot_data <- tibble(
  Fitted = fitted(model1.2_nb_SSB),
  Residuals = residuals(model1.2_nb_SSB))

# Plot the residuals vs fitted values
ggplot(plot_data, aes(x = Fitted, y = Residuals)) +
  geom_point() +
  theme_classic()
```

#### Simulate Residuals with DHARMa

```{r simulate residuals}

# Simulate residuals 
model1.2_nb_SSB_simulation <- simulateResiduals(fittedModel = model1.2_nb_SSB)

# Plot simulated residuals
plot(model1.2_nb_SSB_simulation)
```

Simulated residuals deviate significantly from expectations, are overdispersed and have many outliers. The within-group deviations from uniformity are significant. Test for homogeneity of variance is significant.

#### Test for Overdispersion

```{r test for overdispersion}
testDispersion(model1.2_nb_SSB_simulation)
```

Simulated residuals significantly overdispersed.

#### Test for Zero-Inflation

```{r test for zero inflation}
testZeroInflation(model1.2_nb_SSB_simulation)
```

Simulated residuals significantly zero inflated.

## Model 1.3.

### Model Validation

#### Convergence Check

```{r convergence check}

# Load model from file
load(here::here("data", "models", "model1.3_nb_municipality.RData"))

# Get summary of the model
summary(model1.3_nb_municipality)
```

Model output suggests a significant effect of time and a significant interaction between time period and land cover change (Y) but no significant effect of land cover. AIC and BIC values still large. Dispersion parameter = 45.5.

#### Check Residuals

```{r model 1, attempt 3 check residuals}

# Create tibble for plotting
plot_data <- tibble(
  Fitted = fitted(model1.3_nb_municipality),
  Residuals = residuals(model1.3_nb_municipality))

# Plot the residuals vs fitted values
ggplot(plot_data, aes(x = Fitted, y = Residuals)) +
  geom_point() +
  theme_classic()
```

#### Simulate Residuals with DHARMa

```{r model 1 attempt 3 simulate residuals}

# Simulate residuals 
model1.3_nb_municipality_simulation <- simulateResiduals(fittedModel = model1.3_nb_municipality)

# Plot simulated residuals
plot(model1.3_nb_municipality_simulation)
```

#### Test for Overdispersion

```{r model 1 attempt 3 test overdispersion}
testDispersion(model1.3_nb_municipality_simulation)
```

Residuals significantly overdispersed.

#### Test for Zero-Inflation

```{r model 1 attempt 3 test zero-inflation}
testZeroInflation(model1.3_nb_municipality_simulation)
```

Residuals significantly zero-inflated.

## Model 1.4.

### Model Validation

#### Convergence Check

```{r model 1, attempt 4, convergence check}

# Load model from file
load(here::here("data", "models", "model1.4_nb_SSB.RData"))

# Get summary of the model
summary(model1.4_nb_SSB)
```

No significant effect of time or land cover change. Estimates of the effect sizes are also very small.

#### Check Residuals

```{r model 1, attempt 4, check residuals}

# Create tibble for plotting
plot_data <- tibble(
  Fitted = fitted(model1.4_nb_SSB),
  Residuals = residuals(model1.4_nb_SSB))

# Plot the residuals vs fitted values
ggplot(plot_data, aes(x = Fitted, y = Residuals)) +
  geom_point() +
  theme_classic()
```

#### Simulate Residuals

```{r model 1, attempt 4, simulate residuals}

# Simulate residuals 
model1.4_nb_SSB_simulation <- simulateResiduals(fittedModel = model1.4_nb_SSB)

# Plot simulated residuals
plot(model1.4_nb_SSB_simulation)
```

#### Test Overdispersion

```{r model 1, attempt 4, test overdispersion}
testDispersion(model1.4_nb_SSB_simulation)
```

Residuals signficantly overdispersed.

#### Test Zero-Inflation

```{r model 1, attempt 4, test zero-inflation}
testZeroInflation(model1.4_nb_SSB_simulation)
```

Residuals significantly zero-inflated.

## Model 1.5.

### Model Validation

#### Convergence Check

```{r model 1, attempt 5, convergence check}

# Load model from file
load(here::here("data", "models", "model1.5_nb_municipality.RData"))

# Get summary of the model
summary(model1.5_nb_municipality)
```

#### Check Residuals

```{r model 1, attempt 5, check residuals}

# Create tibble for plotting
plot_data <- tibble(
  Fitted = fitted(model1.5_nb_municipality),
  Residuals = residuals(model1.5_nb_municipality))

# Plot the residuals vs fitted values
ggplot(plot_data, aes(x = Fitted, y = Residuals)) +
  geom_point() +
  theme_classic()
```

#### Simulate Residuals

```{r model 1, attempt 5, simulate residuals}

# Simulate residuals 
model1.5_nb_municipality_simulation <- simulateResiduals(fittedModel = model1.5_nb_municipality)

# Plot simulated residuals
plot(model1.5_nb_municipality_simulation)
```

#### Test Overdispersion

```{r model 1, attempt 5, check overdispersion}
testDispersion(model1.5_nb_municipality_simulation) 
```

#### Test Zero-Inflation

```{r model 1, attempt 5, check zero-inflation}
testZeroInflation(model1.5_nb_municipality_simulation) 
```
