---
title: "Land Cover Change (Y/N) and Occurrence Records, Time and Offset"
author: "Beatrice Trascau"
date: "2024-06-11"
output: html_document
---

# **Introduction**

This markdown outlines the process of modelling the number of occurrence records after a land cover change in relation to land cover changes, time and SSB ID with the number of occurrences before the land cover changes as an offset.

The models ran as part of this are as follows:

| Model Number | Function | Family  | Link | SSBID/Municipality | Ran?                                            |
|--------------|----------|---------|------|--------------------|-------------------------------------------------|
| 2.1          | glmer.nb |         |      | SSBID              | No - too large                                  |
| 2.2          | glmer.nb |         |      | SSBID              | No - NaNs introduced                            |
| 2.3.         | glmmTMB  | nbinom1 | log  | SSBID              | No - negative log-likeligood is NaN at starting |
| 2.4          | glmmTMB  | nbinom1 | log  | Municipality       | No - negative log-likeligood is NaN at starting |
| 2.5          | glmmTMB  | nbinom2 | log  | SSBID              | YES!                                            |

```{r library load, include = FALSE}
library(here)
library(dplyr)
library(lattice)
library(lme4)
library(DHARMa)
library(glmmTMB)
library(mgcv)
library(gamm4)
library(ggplot2)
```

# Data Exploration

Inspect the dataframe used for modeling that also contains the number of occurrences before a land cover change.

## Data Exploration

### Inspect Data

```{r inspect before after data}

# Read in data
load(here::here("data", "derived_data",
                "occ_y_n_cover_change_before_after_for_modell.rda"))

glimpse(occ_y_n_cover_change_before_after_for_modell)
```

### Check for Outliers

```{r check outliers}
# Barplot of number of occurrences
# ggplot(occ_y_n_cover_change_before_after_for_modell , 
#        aes(x = cover_change, y = ocurrences_before, fill = time_period))+
#   geom_boxplot()+
#   theme_classic()
```

Just as with the occurrences after the land cover change, the occurrences before also have some very high outliers, more so for the pixels where there is no change in land cover.

### Check Normality

```{r normality before}

# Plot histogram of number of records by time period
# occ_y_n_cover_change_before_after_for_modell |>
#   filter(ocurrences_before < 100) |>
#   ggplot(aes(x = ocurrences_before)) +
#   geom_histogram(binwidth = 2, position = "dodge", 
#                  alpha = 0.4) +
#   facet_grid(time_period ~ ., scales = "free") +
#   theme_classic()
```

Distribution not normal.

### Check Zero Inflation

```{r zero inflation before after}

# Calculate % of rows with values of 0 for number of occurrences
100 * sum(occ_y_n_cover_change_before_after_for_modell$ocurrences_after == 0) / nrow(occ_y_n_cover_change_before_after_for_modell) #0.001

100 * sum(occ_y_n_cover_change_before_after_for_modell$ocurrences_before == 0) / nrow(occ_y_n_cover_change_before_after_for_modell) # 0.007

```

No problems with zero inflation - or so it seems.

### Check Relationships

```{r check relationship before after}
# occ_y_n_cover_change_before_after_for_modell |>
#   ggplot(aes(x = ocurrences_before, y = ocurrences_after, fill = cover_change)) +
#   geom_point(size = 2) +
#   facet_wrap(~time_period) +
#   theme_classic()
```

# Model 2.1.

ocurrences_after \~ cover_change \* time_period \* ocurrences_before + offset(ocurrences_before) + (1 \| SSBID)

Error in qr.default(X, tol = tol, LAPACK = FALSE) : too large a matrix for LINPACK.

This error seems to be related to running out of memory to run the model with. I will try to run the models on a random subset of the data.

# Model 2.2.

This is identical to model 2.1. but on a subset of the data.

Error in eval(mc, parent.frame(1L)) : PIRLS loop resulted in NaN.

# Model 2.3. & 2.4.

ocurrences_after \~ cover_change \* time_period \* ocurrences_before + offset(ocurrences_before) + (1 \| SSBID)

Model did not run: Error in fitTMB(TMBStruc) : negative log-likelihood is NaN at starting parameter values. This is likley caused by the extreme values in the number of occurrences.

# Model 2.5.

## Model Validation

### Convergence Check

```{r model 2.5. convergence check}

# Load model from file
load(here::here("data", "models", "model2.5_nb.RData"))

# Get summary of the model
summary(model2.5_nb)

```

Significant effect of time, the occurrences before, cover change + time, and cover change + time + occurrences before.

### Check Residuals

```{r model 2.5. check residuals}
# Extract residuals and fitted values
# residuals <- residuals(model2.5_nb, type = "pearson")
# fitted <- fitted(model2.5_nb)
# 
# # Plot residuals vs fitted values
# plot(fitted, residuals)
# abline(h = 0, col = "red")
# title("Residuals vs Fitted")
# 
# # Histogram of residuals
# hist(residuals, breaks = 30, main = "Histogram of Residuals", xlab = "Residuals")
# 
# # Q-Q plot
# qqnorm(residuals)
# qqline(residuals, col = "red")
```

```{r model 2.5. plot residuals vs fitted}

# Create tibble for plotting
plot_data <- tibble(
  Fitted = fitted(model2.5_nb),
  Residuals = residuals(model2.5_nb))

# Plot the residuals vs fitted values
ggplot(plot_data, aes(x = Fitted, y = Residuals)) +
  geom_point() +
  theme_classic()
```

### Check Random Effects

```{r model 2.5. check random effects}
#ranef(model2.5_nb)
```

### Simulate Residuals with DHARMa

```{r model 2.5. simulate residuals}

# Simulate residuals 
#model2.5_nb_SSB_simulation <- simulateResiduals(fittedModel = model2.5_nb)

# Plot simulated residuals
#plot(model2.5_nb_SSB_simulation)
```

### Test for Overdispersion with DHARMa

```{r model 2.5. check overdispersion}
#testDispersion(model2.5_nb_SSB_simulation)
```

### Test Overdispersion manually

```{r model 2.5. manual overdispersion check}

# # Get residuals
# residuals <- residuals(model2.5_nb)
# 
# # Compute residual deviance and degrees of freedom
# res_deviance <- sum(residuals^2)
# df <- nrow(subset_data_intens_extens_before_after) - length(fixef(model4.1)$cond)
# 
# # Overdispersion test
# overdispersion <- res_deviance / df
# overdispersion

```

### Test for Zero-Inflation

```{r model 2.5. check zero inflation}
#testZeroInflation(model2.5_nb_SSB_simulation)
```

Simulating residuals is too memory intensive.
