---
title: "glmms_land_cover_change(intens/extens)_and_occurrences"
author: "Beatrice Trascau"
date: "2024-06-09"
output: html_document
---

# **3.2. GLMMS: Land Cover Change (Intensification/Extensification) and Occurrence Records**

This markdown outlines the process of modelling the number of species occurrence records in relation to land cover changes, time and SSB ID. Here, land cover changes are categorised as representing either a form of intensification or extensification or land-use.

```{r library load, include = FALSE}
library(here)
library(dplyr)
library(lattice)
library(lme4)
library(DHARMa)
library(glmmTMB)
library(mgcv)
library(gamm4)
library(ggplot2)

```

```{r data load, include = FALSE, eval = FALSE}
load(here("data", "derived_data", "occurrences_SSB_municipalities_land_cover.rda"))
occ_SSB_land_cover <- occurrence_municipalities_df
```

## Data Preparations for Model

```{r prepare df for analyis, include = FALSE, eval = FALSE}

# 1. First period of change: 2000-2006 -------------------------------------------------

# Change column names for easier df manipulation
occ_SSB_land_cover <- occ_SSB_land_cover |>
  rename(land_cover2000 = U2006_CLC2000_V2020_20u1,
         land_cover2006 = U2012_CLC2006_V2020_20u1,
         land_cover2012 = U2018_CLC2012_V2020_20u1,
         land_cover2018 = U2018_CLC2018_V2020_20u1)

# Prepare data for the period 2006-2009
occ_df_after_2000.2006 <- occ_SSB_land_cover |>
  select(V1, gbifID, year, species, land_cover2000, land_cover2006, 
         SSBID, cell_ID, NAME_2) |>
  filter(!is.na(land_cover2000) & !is.na(land_cover2006)) |>
  filter(year >= 2006 & year <= 2009) |>
  mutate(cover_change = case_when(
    land_cover2000 - land_cover2006 %in% c(79, 102, 249, 379, 589, 710,
                                           23, 147, 170, 608, 631, -102,
                                           -79, 487, 510, 277, 300, -340,
                                           -331, -461) ~ "Intensification",
    land_cover2000 - land_cover2006 %in% c(-249, -379, -589, -170, -300, -510,
                                           -147, -277, -487, 130, -210, 461, 331,
                                           121, 340, -23, -710, -121, -608) ~ "Extensification",
    land_cover2000 - land_cover2006 == 0 ~ "No_change"
  ))

# Calculate number of records for the period 2006-2009
occ_df_after_2000.2006_records <- occ_df_after_2000.2006 |>
  group_by(cell_ID) |>
  summarise(
    ocurrences_after = n(),
    land_cover2000 = first(land_cover2000),
    land_cover2006 = first(land_cover2006),
    cover_change = first(cover_change),
    SSBID = first(SSBID),
    NAME_2 = first(NAME_2)) |>
  mutate(time_period = "2000-2006") |>
  select(-c(land_cover2000, land_cover2006))

# 2. Second period of change: 2006 - 2012 ----------------------------------------------

# Prepare data for the period 2012-2015
occ_df_after_2006.2012 <- occ_SSB_land_cover |>
  select(V1, gbifID, year, species, land_cover2006, land_cover2012, 
         SSBID, cell_ID, NAME_2) |>
  filter(!is.na(land_cover2006) & !is.na(land_cover2012)) |>
  filter(year >= 2012 & year <= 2015) |>
  mutate(cover_change = case_when(
    land_cover2000 - land_cover2006 %in% c(79, 102, 249, 379, 589, 710,
                                           23, 147, 170, 608, 631, -102,
                                           -79, 487, 510, 277, 300, -340,
                                           -331, -461) ~ "Intensification",
    land_cover2000 - land_cover2006 %in% c(-249, -379, -589, -170, -300, -510,
                                           -147, -277, -487, 130, -210, 461, 331,
                                           121, 340, -23, -710, -121, -608) ~ "Extensification",
    land_cover2000 - land_cover2006 == 0 ~ "No_change"
  ))

# Calculate number of records for the period 2012-2015
occ_df_after_2006.2012_records <- occ_df_after_2006.2012 |>
  group_by(cell_ID) |>
  summarise(
    ocurrences_after = n(),
    land_cover2006 = first(land_cover2006),
    land_cover2012 = first(land_cover2012),
    cover_change = first(cover_change),
    SSBID = first(SSBID),
    NAME_2 = first(NAME_2)) |>
  mutate(time_period = "2006-2012") |>
  select(-c(land_cover2006, land_cover2012))

# 3. Third period of change: 2012 - 2018 ----------------------------------------------

# Prepare data for the period 2015-2018
occ_df_after_2012.2018 <- occ_SSB_land_cover |>
  select(V1, gbifID, year, species, land_cover2012, land_cover2018, 
         SSBID, cell_ID, NAME_2) |>
  filter(!is.na(land_cover2012) & !is.na(land_cover2018)) |>
  filter(year >= 2015 & year <= 2018) |>
  mutate(cover_change = case_when(
    land_cover2000 - land_cover2006 %in% c(79, 102, 249, 379, 589, 710,
                                           23, 147, 170, 608, 631, -102,
                                           -79, 487, 510, 277, 300, -340,
                                           -331, -461) ~ "Intensification",
    land_cover2000 - land_cover2006 %in% c(-249, -379, -589, -170, -300, -510,
                                           -147, -277, -487, 130, -210, 461, 331,
                                           121, 340, -23, -710, -121, -608) ~ "Extensification",
    land_cover2000 - land_cover2006 == 0 ~ "No_change"
  ))

# Calculate number of records for the period 1997-2000
occ_df_after_2012.2018_records <- occ_df_after_2012.2018 |>
  group_by(cell_ID) |>
  summarise(
    ocurrences_after = n(),
    land_cover2012 = first(land_cover2012),
    land_cover2018 = first(land_cover2018),
    cover_change = first(cover_change),
    SSBID = first(SSBID),
    NAME_2 = first(NAME_2)) |>
  mutate(time_period = "2012-2018") |>
  select(-c(land_cover2012, land_cover2018))

# 4. Combine the above 3 into a single df ----------------------------------------------

occ_intens_extens_after_records_for_model <- bind_rows(occ_df_after_2000.2006_records,
                                  occ_df_after_2006.2012_records,
                                  occ_df_after_2012.2018_records) |>
  select(-cell_ID) |>
  rename(municipality = NAME_2)

# 5. Write dataframe to file -----------------------------------------------------------
save(occ_intens_extens_after_records_for_model, file = here::here("data", "derived_data",
                                          "occ_intens_extens_after_records_for_model.rda"))

```

## Data Exploration

### Inspect Data

```{r inspect data}
# Read data in 

glimpse(occ_intens_extens_after_records_for_model)

```

### Check for Outliers

```{r check outliers}

# Barplot of number of occurrences
ggplot(occ_intens_extens_after_records_for_model , 
       aes(x = cover_change, y = ocurrences_after, fill = time_period))+
  geom_boxplot()+
  theme_classic()
```

### Check Normality

```{r check normality}

# Plot histogram of number of records by time period
occ_intens_extens_after_records_for_model |>
  filter(ocurrences_after < 100) |>
  ggplot(aes(x = ocurrences_after)) +
  geom_histogram(binwidth = 2, position = "dodge", 
                 alpha = 0.4) +
  facet_grid(time_period ~ ., scales = "free") +
  theme_classic()
```

### Check Zero Inflation

```{r check zero inflation}

# Calculate % of rows with values of 0 for number of occurrences
100 * sum(occ_intens_extens_after_records_for_model$ocurrences_after == 0) / nrow(occ_intens_extens_after_records_for_model)
#suspicious
```

### Check for Collinearity

Covariates in model (cover change, SSB ID and time period) are all categorial. No need to check for collinearity between them.

### Check Relationships

```{r check relationships}

# Plot relationships
occ_intens_extens_after_records_for_model |>
  ggplot(aes(x = time_period, y = ocurrences_after)) +
  geom_point() +
  facet_wrap(~cover_change) +
  theme_classic()

```

## Model 3: Attempt 1: Number of occurrences after change by land cover, time period and SSB ID - NB glmmTMB

```{r model 3 attempt 1 include = FALSE, eval = FALSE}

# Run negative binomial model
model3_extens_intens_nb_SSB <- glmmTMB(ocurrences_after ~ cover_change * time_period + (1|SSBID),
                          family = nbinom1, 
                          data = occ_intens_extens_after_records_for_model)

# Save model output to file to save time next time
save(model3_extens_intens_nb_SSB, file = here::here("data", "models", "model3_extens_intens_nb_SSB.RData"))
```
