---
title: "Land Cover Change (Intensification/Extensification) and Occurrence Records & Time"
author: "Beatrice Trascau"
date: "2024-06-09"
output: html_document
---

# **Introduction**

This markdown outlines the process of modelling the number of species occurrence records in relation to land cover changes, time and SSB ID. Here, land cover changes are categorised as representing either a form of intensification or extensification or land-use. These models were ran on a 10% subset of the data (\~3 mil records) to save computational power.

The defined model was:

occurrences_after \~ cover_change (intens/extens/no change) \* time_period + (1\|SSBID)

occurrences_after \~ cover_change (intens/extens/no change) \* time_period + (1\|municipality)

The models variants ran as part of this are as follows:

| Model Number | Function | Family  | Link  | SSBID/Municipality | Ran? |
|--------------|----------|---------|-------|--------------------|------|
| 3.1.         | glmmTMB  | nbinom1 | log   | SSBID              |      |
| 3.2.         | glmmTMB  | nbinom1 | log   | Municipality       |      |
| 3.3.         | glmmTMB  | nbinom1 | logit | SSBID              |      |
| 3.4.         | glmmTMB  | nbinom1 | logit | Municipality       |      |
| 3.5.         | glmmTMB  | nbinom2 | log   | SSBID              |      |
| 3.6.         | glmmTMB  | nbinom2 | log   | Municipality       |      |
| 3.7.         | glmmTMB  | nbinom2 | logit | SSBID              |      |
| 3.8.         | glmmTMB  | nbinom2 | logit | Municipality       |      |
|              |          |         |       |                    |      |

```{r library load, include = FALSE}
library(here)
library(dplyr)
library(lattice)
library(lme4)
library(DHARMa)
library(glmmTMB)
library(mgcv)
library(gamm4)
library(ggplot2)

```

# Data Exploration

## Inspect Data

```{r inspect data}

# Read data in 
load(here::here("data", "derived_data",
                "occ_intens_extens_after_records_for_model.rda"))

# Check data structure
glimpse(occ_intens_extens_after_records_for_model)

```

## Check for Outliers

```{r check outliers}

# Barplot of number of occurrences
ggplot(occ_intens_extens_after_records_for_model , 
       aes(x = cover_change, y = ocurrences_after, fill = time_period))+
  geom_boxplot()+
  theme_classic()
```

## Check Normality

```{r check normality}

# Plot histogram of number of records by time period
occ_intens_extens_after_records_for_model |>
  filter(ocurrences_after < 100) |>
  ggplot(aes(x = ocurrences_after)) +
  geom_histogram(binwidth = 2, position = "dodge", 
                 alpha = 0.4) +
  facet_grid(time_period ~ ., scales = "free") +
  theme_classic()
```

## Check Zero Inflation

```{r check zero inflation}

# Calculate % of rows with values of 0 for number of occurrences
100 * sum(occ_intens_extens_after_records_for_model$ocurrences_after == 0) / nrow(occ_intens_extens_after_records_for_model)
#suspicious
```

## Check for Collinearity

Covariates in model (cover change, SSB ID and time period) are all categorial. No need to check for collinearity between them.

## Check Relationships

```{r check relationships}

# Plot relationships
occ_intens_extens_after_records_for_model |>
  ggplot(aes(x = time_period, y = ocurrences_after)) +
  geom_point() +
  facet_wrap(~cover_change) +
  theme_classic()

```

# Model 3.1.

```{r model 3.1. convergence check}
# Load model from file
load(here::here("data", "models", "model3_extens_intens_nb_SSB.RData"))

# Get summary of the model
summary(model3_extens_intens_nb_SSB)
```

## Check Residuals

```{r model 3.1. check residuals}

# Create tibble for plotting
plot_data <- tibble(
  Fitted = fitted(model3_extens_intens_nb_SSB),
  Residuals = residuals(model3_extens_intens_nb_SSB))

# Plot the residuals vs fitted values
ggplot(plot_data, aes(x = Fitted, y = Residuals)) +
  geom_point() +
  theme_classic()
```

# Model 3.2.

```{r model 3.2. convergence check}
# Load model from file
load(here::here("data", "models", "model3.2_extens_intens_nb_SSB.RData"))

# Get summary of the model
summary(model3.2_extens_intens_nb_SSB)
```

## Check Residuals

```{r model 3.2. check residuals}

# Create tibble for plotting
plot_data <- tibble(
  Fitted = fitted(model3.2_extens_intens_nb_SSB),
  Residuals = residuals(model3.2_extens_intens_nb_SSB))

# Plot the residuals vs fitted values
ggplot(plot_data, aes(x = Fitted, y = Residuals)) +
  geom_point() +
  theme_classic()
```

# Model 3.3.

```{r model 3.3. convergence check}
# Load model from file
load(here::here("data", "models", "model3.3_extens_intens_nb_SSB.RData"))

# Get summary of the model
summary(model3.3_extens_intens_nb_SSB)
```

## Check Residuals

```{r model 3.3. check residuals}

# Create tibble for plotting
plot_data <- tibble(
  Fitted = fitted(model3.3_extens_intens_nb_SSB),
  Residuals = residuals(model3.3_extens_intens_nb_SSB))

# Plot the residuals vs fitted values
ggplot(plot_data, aes(x = Fitted, y = Residuals)) +
  geom_point() +
  theme_classic()
```

# Model 3.4.

```{r model 3.4. convergence check}
# Load model from file
load(here::here("data", "models", "model3.4_extens_intens_nb_SSB.RData"))

# Get summary of the model
summary(model3.4_extens_intens_nb_SSB)
```

## Check Residuals

```{r model 3.4. check residuals}

# Create tibble for plotting
plot_data <- tibble(
  Fitted = fitted(model3.4_extens_intens_nb_SSB),
  Residuals = residuals(model3.4_extens_intens_nb_SSB))

# Plot the residuals vs fitted values
ggplot(plot_data, aes(x = Fitted, y = Residuals)) +
  geom_point() +
  theme_classic()
```

# Model 3.5.

```{r model 3.5. convergence check}
# Load model from file
load(here::here("data", "models", "model3.5_extens_intens_nb_SSB.RData"))

# Get summary of the model
summary(model3.5_extens_intens_nb_SSB)
```

## Check Residuals

```{r model 3.5. check residuals}

# Create tibble for plotting
plot_data <- tibble(
  Fitted = fitted(model3.5_extens_intens_nb_SSB),
  Residuals = residuals(model3.5_extens_intens_nb_SSB))

# Plot the residuals vs fitted values
ggplot(plot_data, aes(x = Fitted, y = Residuals)) +
  geom_point() +
  theme_classic()
```

# Model 3.6.

```{r model 3.6. convergence check}
# Load model from file
load(here::here("data", "models", "model3.6_extens_intens_nb_SSB.RData"))

# Get summary of the model
summary(model3.6_extens_intens_nb_SSB)
```

## Check Residuals

```{r model 3.6. check residuals}

# Create tibble for plotting
plot_data <- tibble(
  Fitted = fitted(model3.6_extens_intens_nb_SSB),
  Residuals = residuals(model3.6_extens_intens_nb_SSB))

# Plot the residuals vs fitted values
ggplot(plot_data, aes(x = Fitted, y = Residuals)) +
  geom_point() +
  theme_classic()
```

# Model 3.7.

```{r model 3.7. convergence check}
# Load model from file
load(here::here("data", "models", "model3.7_extens_intens_nb_SSB.RData"))

# Get summary of the model
summary(model3.7_extens_intens_nb_SSB)
```

## Check Residuals

```{r model 3.7. check residuals}

# Create tibble for plotting
plot_data <- tibble(
  Fitted = fitted(model3.7_extens_intens_nb_SSB),
  Residuals = residuals(model3.7_extens_intens_nb_SSB))

# Plot the residuals vs fitted values
ggplot(plot_data, aes(x = Fitted, y = Residuals)) +
  geom_point() +
  theme_classic()
```

# Model 3.8.

```{r model 3.8. convergence check}
# Load model from file
load(here::here("data", "models", "model3.8_extens_intens_nb_SSB.RData"))

# Get summary of the model
summary(model3.8_extens_intens_nb_SSB)
```

## Check Residuals

```{r model 3.8. check residuals}

# Create tibble for plotting
plot_data <- tibble(
  Fitted = fitted(model3.8_extens_intens_nb_SSB),
  Residuals = residuals(model3.8_extens_intens_nb_SSB))

# Plot the residuals vs fitted values
ggplot(plot_data, aes(x = Fitted, y = Residuals)) +
  geom_point() +
  theme_classic()
```
