---
title: "Land Cover Change (Intensification/Extensification) and Occurrence Records, Time and Offset"
author: "Beatrice Trascau"
date: "2024-06-12"
output: html_document
---

# Introduction

This markdown outlines the process of modelling the number of occurrence records after a land cover change in relation to land cover changes, time and SSB ID with the number of occurrences before the land cover changes as an offset. Here, land cover changes are categorised as representing either a form of intensification or extensification or land-use. These models were ran on a 10% subset of the data (\~3 mil records) to save computational power.

The defined model was:

occurrences_after \~ cover_change (intens/extens/no change) \* time_period \* occurrence_before + offset(occurrence_before) + (1\|SSBID) (Model 4.1.)

occurrences_after \~ cover_change (intens/extens/no change) \* time_period \* occurrence_before + offset(occurrence_before) + (1\|municipality) (Model 4.2.)

Both models were ran with the glmmTMB package, family nbinom2, log link.

```{r library load, include = FALSE}
library(here)
library(dplyr)
library(lattice)
library(lme4)
library(DHARMa)
library(glmmTMB)
library(mgcv)
library(gamm4)
library(ggplot2)
```

# Data Exploration

Inspect the dataframe used for modeling that also contains the number of occurrences before a land cover change.

## Data Exploration

### Inspect Data

```{r inspect before after data}

# Read in data
load(here::here("data", "derived_data",
                "occ_intens_extens_before_after_for_model.rda"))

# Get glimpse of data
glimpse(occ_intens_extens_before_after_for_model)

```

Size of data does not allow easy data exploration.

# Model 4.1.

## Model Validation

### Convergence Check

```{r model 4.1. convergence check}

# Load model from file
load(here::here("data", "models", "model4.1.RData"))

# Get summary of the model
summary(model4.1)

```

### Check Residuals

```{r model 4.1. check residuals}

# Create tibble for plotting
plot_data <- tibble(
  Fitted = fitted(model4.1),
  Residuals = residuals(model4.1))

# Plot the residuals vs fitted values
ggplot(plot_data, aes(x = Fitted, y = Residuals)) +
  geom_point() +
  theme_classic()
```

# Model 4.2.

## Model Validation

### Convergence Check

```{r model 4.2. convergence check}

# Load model from file
load(here::here("data", "models", "model4.2.RData"))

# Get summary of the model
summary(model4.2)

```

### Check Residuals

```{r model 4.2. check residuals}

# Create tibble for plotting
plot_data <- tibble(
  Fitted = fitted(model4.2),
  Residuals = residuals(model4.2))

# Plot the residuals vs fitted values
ggplot(plot_data, aes(x = Fitted, y = Residuals)) +
  geom_point() +
  theme_classic()
```
