---
title: "Land Cover Change (Types) and Occurrence Records & Time"
author: "Beatrice Trascau"
date: "2024-06-12"
output: html_document
---

# Introduction

This markdown outlines the process of modelling the number of species occurrence records in relation to land cover changes, time and SSB ID (or Municipality). Here, we use the individual types of land cover changes. These models were ran on a 10% subset of the data to save computational power.

The defined model was:

-   occurrences_after \~ cover_change(types) \* time_period + (1\|SSBID) (model3.1 & model3.3)

-   occurrences_after \~ cover_change(types) \* time_period + (1\|municipality) (model3.2 & model3.4)

-   occurrences_after \~ cover_change(types) \* time_period \* occurrences_before + offset(occurrences_before) + (1\|SSBID) (model3.5)

-   occurrences_after \~ cover_change(types) \* time_period \* occurrences_before + offset(occurrences_before) + (1\|municipality) (model3.6)

The models variants ran as part of this are as follows:

| Model Number | Function | Family  | Link | SSBID/Municipality | Offset? |
|--------------|----------|---------|------|--------------------|---------|
| 3.1.         | glmmTMB  | nbinom1 | log  | SSBID              | N       |
| 3.2.         | glmmTMB  | nbinom1 | log  | Municipality       | N       |
| 3.3.         | glmmTMB  | nbinom2 | log  | SSBID              | N       |
| 3.4.         | glmmTMB  | nbinom2 | log  | Municipality       | N       |
| 3.5.         | glmmTMB  | nbinom2 | log  | SSBID              | Y       |
| 3.6.         | glmmTMB  | nbinom2 | log  | Municipality       | Y       |

```{r library load, include = FALSE}
library(here)
library(dplyr)
library(DHARMa)
library(ggplot2)
```

# Data Exploration

## Inspect Data

```{r inspect data}

# Read data in 
load(here::here("data", "derived_data",
                "occ_cover_change_types_before_after_for_model.rda"))

# Check data structure
glimpse(occ_cover_change_types_before_after_for_model)
```

## Check for Outliers

```{r check outliers}

# Barplot of number of occurrences
ggplot(occ_cover_change_types_before_after_for_model , 
       aes(x = cover_change, y = ocurrences_after, fill = time_period))+
  geom_boxplot()+
  theme_classic()
```

## Check Normality

```{r check normality}

# Plot histogram of number of records by time period
occ_cover_change_types_before_after_for_model |>
  filter(ocurrences_after < 100) |>
  ggplot(aes(x = ocurrences_after)) +
  geom_histogram(binwidth = 2, position = "dodge", 
                 alpha = 0.4) +
  facet_grid(time_period ~ ., scales = "free") +
  theme_classic()
```

## Check Relationships

```{r check relationships}

# Plot relationships
occ_cover_change_types_before_after_for_model |>
  ggplot(aes(x = time_period, y = ocurrences_after)) +
  geom_point() +
  facet_wrap(~cover_change) +
  theme_classic()
```

# Model 3.1.

```{r model 3.1. convergence check}
# Load model from file
load(here::here("data", "models", "model3.1_SSB.RData"))

# Get summary of the model
summary(model3.1_SSB)
```

## Check Residuals

```{r model 3.1. check residuals}

# Create tibble for plotting
plot_data <- tibble(
  Fitted = fitted(model3.1_SSB),
  Residuals = residuals(model3.1_SSB))

# Plot the residuals vs fitted values
ggplot(plot_data, aes(x = Fitted, y = Residuals)) +
  geom_point() +
  theme_classic()
```

# Model 3.2.

```{r model 3.2. convergence check}
# Load model from file
load(here::here("data", "models", "model3.2_municipality.RData"))

# Get summary of the model
summary(model3.2_municipality)
```

## Check Residuals

```{r model 3.2. check residuals}

# Create tibble for plotting
plot_data <- tibble(
  Fitted = fitted(model3.2_municipality),
  Residuals = residuals(model3.2_municipality))

# Plot the residuals vs fitted values
ggplot(plot_data, aes(x = Fitted, y = Residuals)) +
  geom_point() +
  theme_classic()
```

# Model 5.3.

```{r model 3.3. convergence check}
# Load model from file
load(here::here("data", "models", "model3.3_SSB.RData"))

# Get summary of the model
summary(model3.3_SSB)
```

## Check Residuals

```{r model 3.3. check residuals}

# Create tibble for plotting
plot_data <- tibble(
  Fitted = fitted(model3.3_SSB),
  Residuals = residuals(model3.3_SSB))

# Plot the residuals vs fitted values
ggplot(plot_data, aes(x = Fitted, y = Residuals)) +
  geom_point() +
  theme_classic()
```

# Model 3.4.

```{r model 3.4. convergence check}
# Load model from file
load(here::here("data", "models", "model3.4_municipality.RData"))

# Get summary of the model
summary(model3.4_municipality)
```

## Check Residuals

```{r model 3.4. check residuals}

# Create tibble for plotting
plot_data <- tibble(
  Fitted = fitted(model3.4_municipality),
  Residuals = residuals(model3.4_municipality))

# Plot the residuals vs fitted values
ggplot(plot_data, aes(x = Fitted, y = Residuals)) +
  geom_point() +
  theme_classic()
```

# Model 3.5.

```{r model 3.5. convergence check}
# Load model from file
load(here::here("data", "models", "model3.5_SSB.RData"))

# Get summary of the model
summary(model3.5_SSB)
```

## Check Residuals

```{r model 3.5. check residuals}

# Create tibble for plotting
plot_data <- tibble(
  Fitted = fitted(model3.5_SSB),
  Residuals = residuals(model3.5_SSB))

# Plot the residuals vs fitted values
ggplot(plot_data, aes(x = Fitted, y = Residuals)) +
  geom_point() +
  theme_classic()
```

# Model 3.6.

```{r model 3.6. convergence check}
# Load model from file
load(here::here("data", "models", "model3.6_municipality.RData"))

# Get summary of the model
summary(model3.6_municipality)
```

## Check Residuals

```{r model 3.6. check residuals}

# Create tibble for plotting
plot_data <- tibble(
  Fitted = fitted(model3.6_municipality),
  Residuals = residuals(model3.6_municipality))

# Plot the residuals vs fitted values
ggplot(plot_data, aes(x = Fitted, y = Residuals)) +
  geom_point() +
  theme_classic()
```
