---
title: "Land Cover Change (Y/N) and Occurrence Records & Time"
author: "Beatrice Trascau"
date: "2025-02-25"
output: html_document
---

```{r}
library(here)
library(dplyr)
library(DHARMa)
library(ggplot2)
library(knitr)
library(bbmle)
```

# Introduction

This markdown outlines the process of modelling the number of occurrence records after a land cover change in relation to land cover changes (Y/N), time and SSB ID, without and with an offset.

The models ran as part of this are as follows:

| Model Name                            | Interaction?                | Marginal Value |
|---------------------------------------|-----------------------------|----------------|
| YN_model1_SSB_interaction             | cover_change \* time_period | 0.001          |
| YN_model2_SSB_no_interaction          | cover_change + time_period  | 0.001          |
| YN_model3_SSB_interaction_0.1_offset  | cover_change \* time_period | 0.1            |
| YN_model4_SSB_interaction_0.01_offset | cover_change \* time_period | 0.01           |

# YN_model1_SSB_interaction

## Get output

```{r YN model 1}

# Load data
load(here("data", "derived_data", "occ_y_n_cover_change_before_after_for_modell.rda"))

# Load model
load(here::here("data", "models", "YN_model1_SSB_interaction.RData"))

# Get model summary
YN_model1_summary <- summary(YN_model1_SSB_interaction)

# Convert summary coefficients to df
YN_model_coeff_df <- as.data.frame(YN_model1_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
YN_model_coeff_df[] <- lapply(YN_model_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(YN_model_coeff_df)
```

# YN_model1_SSB_no_interaction

## Get output

```{r YN model 2}

# Load model
load(here::here("data", "models", "YN_model2_SSB_no_interaction.RData"))

# Get model summary
YN_model2_summary <- summary(YN_model2_SSB_no_interaction)

# Convert summary coefficients to df
YN_model2_coeff_df <- as.data.frame(YN_model2_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
YN_model2_coeff_df[] <- lapply(YN_model2_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(YN_model2_coeff_df)
```

## Compare Model 1 and Model 2

```{r compare AICs }
AICctab(YN_model1_SSB_interaction, YN_model2_SSB_no_interaction, base = TRUE)

#                              AICc      dAICc     df
# YN_model2_SSB_no_interaction 3931684.9       0.0 6 
# YN_model1_SSB_interaction    3931688.1       3.2 8 
```

# Model 3: 0.1 offset

```{r YN model 3}

# Load model
load(here::here("data", "models", "YN_model3_SSB_interaction_0.1_offset.RData"))

# Get model summary
YN_model3_summary <- summary(YN_model3_SSB_interaction_0.1_offset)

# Convert summary coefficients to df
YN_model3_coeff_df <- as.data.frame(YN_model3_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
YN_model3_coeff_df[] <- lapply(YN_model3_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(YN_model3_coeff_df)
```

# Model 4: 0.01 offset

```{r model4}

# Load model
load(here::here("data", "models", "YN_model4_SSB_interaction_0.01_offset.RData"))

# Get model summary
YN_model4_summary <- summary(YN_model4_SSB_interaction_0.01_offset)

# Convert summary coefficients to df
YN_model4_coeff_df <- as.data.frame(YN_model4_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
YN_model4_coeff_df[] <- lapply(YN_model4_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(YN_model4_coeff_df)
```

# Model 5: 0.1 offset and no interaction

```{r YN model 5}

# Load model
load(here::here("data", "models", "YN_model5_SSB_no_interaction_0.1_offset.RData"))

# Get model summary
YN_model5_summary <- summary(YN_model5_SSB_no_interaction_0.1_offset)

# Convert summary coefficients to df
YN_model5_coeff_df <- as.data.frame(YN_model5_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
YN_model5_coeff_df[] <- lapply(YN_model5_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(YN_model5_coeff_df)
```

# AIC Comparison Model 3 and Model 5

```{r model 3 & model 5 AIC}
AICctab(YN_model3_SSB_interaction_0.1_offset, YN_model5_SSB_no_interaction_0.1_offset, base = TRUE)
#                                         AICc      dAICc     df
# YN_model5_SSB_no_interaction_0.1_offset 3589299.7       0.0 6 
# YN_model3_SSB_interaction_0.1_offset    3589302.4       2.7 8 
```

# Model Validation

Doing some model validation on Model 5 (0.1 offset and no interaction).

```{r model 5 residuals}

simulationOutput <- simulateResiduals(fittedModel = YN_model3_SSB_interaction_0.1_offset)

# Plot residual diagnostics
plot(simulationOutput)

# Test specific assumptions
testDispersion(simulationOutput)
testZeroInflation(simulationOutput)
testOutliers(simulationOutput)

```

# Model 6: 0.1 offset, interaction and zero inflation

```{r YN model 6}

# Load model
load(here::here("data", "models", "YN_model6_zero_inflated_interaction_0.1_offset.RData"))

# Get model summary
YN_model6_summary <- summary(YN_ZINB_model)

# Convert summary coefficients to df
YN_model6_coeff_df <- as.data.frame(YN_model6_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
YN_model6_coeff_df[] <- lapply(YN_model6_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(YN_model6_coeff_df)
```

# Model 7: 0.1 offset, no interaction and zero inflation

```{r YN model 7}
# Load model
load(here::here("data", "models", "YN_model7_zero_inflated_no_interaction_0.1_offset.RData"))

# Get model summary
YN_model7_summary <- summary(YN_ZINB_model_no_interaction)

# Convert summary coefficients to df
YN_model7_coeff_df <- as.data.frame(YN_model7_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
YN_model7_coeff_df[] <- lapply(YN_model7_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(YN_model7_coeff_df)
```

# AIC Comparison Model 6 and Model 7

```{r model 6 & 7 AIC}
AICctab(YN_ZINB_model, YN_ZINB_model_no_interaction, base = TRUE)

#                              AICc      dAICc     df
# YN_ZINB_model_no_interaction 3567188.5       0.0 10
# YN_ZINB_model                3567192.3       3.8 12
```
