---
title: "Land Cover Change (Y/N) and Occurrence Records & Time"
author: "Beatrice Trascau"
date: "2024-05-31"
output: html_document
---

```{r}
library(here)
library(dplyr)
library(DHARMa)
library(ggplot2)
```

# Introduction

This markdown outlines the process of modelling the number of occurrence records after a land cover change in relation to land cover changes (Y/N), time and SSB ID, without and with an offset.

The models ran as part of this are as follows:

| Model Number | Function | Family  | Link | SSBID/Municipality | Offset? |
|--------------|----------|---------|------|--------------------|---------|
| 1.1          | glmmTMB  | nbinom1 | log  | SSBID              | N       |
| 1.2          | glmmTMB  | nbinom1 | log  | Municipality       | N       |
| 1.3          | glmmTMB  | nbinom2 | log  | SSBID              | N       |
| 1.4          | glmmTMB  | nbinom2 | log  | Municipality       | N       |
| 1.5          | glmmTMB  | nbinom2 | log  | SSBID              | Y       |
| 1.6          | glmmTMB  | nbinom2 | log  | Municipality       | Y       |

# Data Exploration

## Inspect Data

```{r inspect data}

# Read in data
load(here::here("data", "derived_data",
                "occ_y_n_cover_change_before_after_for_modell.rda"))

glimpse(occ_y_n_cover_change_before_after_for_modell)
```

### Check for Outliers

```{r check outliers}

# Barplot of number of occurrences
ggplot(occ_y_n_cover_change_before_after_for_modell , 
       aes(x = cover_change, y = ocurrences_after, fill = time_period))+
  geom_boxplot()+
  theme_classic()
```

There are some very clear outliers in the number of occurrence records, more so for the pixels where the land cover doesn't change.

### Check Normality

```{r check normality}

# Plot histogram of number of records by time period
occ_y_n_cover_change_before_after_for_modell |>
  filter(ocurrences_after < 100) |>
  ggplot(aes(x = ocurrences_after)) +
  geom_histogram(binwidth = 2, position = "dodge", 
                 alpha = 0.4) +
  facet_grid(time_period ~ ., scales = "free") +
  theme_classic()
```

### Check Zero Inflation

```{r check zero inflation}

# Calculate % of rows with values of 0 for number of occurrences
100 * sum(occ_y_n_cover_change_before_after_for_modell$ocurrences_after == 0) / nrow(occ_y_n_cover_change_before_after_for_modell)
```

### Check for Collinearity

Covariates in model (cover change, SSB ID and time period) are all categorial. No need to check for collinearity between them.

### Check Relationships

```{r check relationships}

# Plot relationships
occ_y_n_cover_change_before_after_for_modell |>
  ggplot(aes(x = time_period, y = ocurrences_after)) +
  geom_point() +
  facet_wrap(~cover_change) +
  theme_classic()
```

Relationships difficult to determine but seems that relationships between the number of occurrences after a land cover change in pixels and time seems to be non-linear.

# Model 1.1.

# Model Validation

### Convergence Check

```{r model 1.1 convergence check}

# Load model from file
load(here::here("data", "models", "model1.1_SSB.RData"))

# Get summary of the model
summary(model1.1_SSB)
```

Model output suggests a significant difference in the number of occurrence records between pixels in 2000-2006 and those in 2006-2012 but no significant effect of land cover change an no significant interaction between time period and land cover change (Y). AIC and BIC values still large. . Dispersion parameter = 26.6.

### Check Residuals

```{r model 1.1 check residuals}

# Create tibble for plotting
plot_data <- tibble(
  Fitted = fitted(model1.1_SSB),
  Residuals = residuals(model1.1_SSB))

# Plot the residuals vs fitted values
ggplot(plot_data, aes(x = Fitted, y = Residuals)) +
  geom_point() +
  theme_classic()
```

### Simulate Residuals with DHARMa

```{r model 1.1 simulate residuals}

# Simulate residuals 
model1.1_SSB_simulation <- simulateResiduals(fittedModel = model1.1_SSB)

# Plot simulated residuals
plot(model1.1_SSB_simulation)
```

Simulated residuals deviate significantly from expectations, are overdispersed and have many outliers. The within-group deviations from uniformity are significant. Test for homogeneity of variance is significant.

### Test for Overdispersion

```{r model 1.1 test for overdispersion}
testDispersion(model1.1_SSB_simulation)
```

Simulated residuals significantly overdispersed.

### Test for Zero-Inflation

```{r model 1.1 test for zero inflation}
testZeroInflation(model1.1_SSB_simulation)
```

Simulated residuals significantly zero inflated.

## Further Validation

```{r}
# Extract residuals and fitted values
residuals <- residuals(model1.1_SSB)
fitted <- fitted(model1.1_SSB)

# Residuals vs. Fitted Values
plot(fitted, residuals, 
     main = "Residuals vs. Fitted Values",
     xlab = "Fitted Values", 
     ylab = "Residuals")
abline(h = 0, col = "red")
```

```{r}
qqnorm(residuals, main = "Q-Q Plot of Residuals")
qqline(residuals, col = "red")
```

# Model 1.2.

## Model Validation

### Convergence Check

```{r model 1.2 convergence check}

# Load model from file
load(here::here("data", "models", "model1.2_municipality.RData"))

# Get summary of the model
summary(model1.2_municipality)
```

Model output suggests a significant difference in the number of occurrence records between pixels in 2000-2006 and those in 2006-2012 but no significant effect of land cover change an no significant interaction between time period and land cover change (Y). AIC and BIC values still large. . Dispersion parameter = 26.6.

### Check Residuals

```{r model 1.2 check residuals}

# Create tibble for plotting
plot_data <- tibble(
  Fitted = fitted(model1.2_municipality),
  Residuals = residuals(model1.2_municipality))

# Plot the residuals vs fitted values
ggplot(plot_data, aes(x = Fitted, y = Residuals)) +
  geom_point() +
  theme_classic()
```

### Simulate Residuals with DHARMa

```{r model 1.2 simulate residuals}

# Simulate residuals 
model1.2_municipality_simulation <- simulateResiduals(fittedModel = model1.2_municipality)

# Plot simulated residuals
plot(model1.2_municipality_simulation)
```

### Test for Overdispersion

```{r model 1.2 test overdispersion}
testDispersion(model1.2_municipality_simulation)
```

Residuals significantly overdispersed.

### Test for Zero-Inflation

```{r model 1.2 test zero-inflation}
testZeroInflation(model1.2_municipality_simulation)
```

Residuals significantly zero-inflated.

# Model 1.3.

## Model Validation

### Convergence Check

```{r model 1.3 convergence check}

# Load model from file
load(here::here("data", "models", "model1.3_SSB.RData"))

# Get summary of the model
summary(model1.3_SSB)
```

Significant effect of time, land cover change and interaction.

### Check Residuals

```{r model 1.3 check residuals}

# Create tibble for plotting
plot_data <- tibble(
  Fitted = fitted(model1.3_SSB),
  Residuals = residuals(model1.3_SSB))

# Plot the residuals vs fitted values
ggplot(plot_data, aes(x = Fitted, y = Residuals)) +
  geom_point() +
  theme_classic()
```

### Simulate Residuals

```{r model 1.3 simulate residuals}

# Simulate residuals 
#model1.3_SSB_simulation <- simulateResiduals(fittedModel = model1.3_SSB)

# Plot simulated residuals
#plot(model1.3_SSB_simulation)
```

### Test Overdispersion

```{r model 1.3, test overdispersion}
#testDispersion(model1.3_SSB_simulation)
```

Residuals signficantly overdispersed.

### Test Zero-Inflation

```{r model 1.3, test zero-inflation}
#testZeroInflation(model1.3_SSB_simulation)
```

Residuals significantly zero-inflated.

# Model 1.4.

## Model Validation

### Convergence Check

```{r model 1.4, convergence check}

# Load model from file
load(here::here("data", "models", "model1.4_municipality.RData"))

# Get summary of the model
summary(model1.4_municipality)
```

There is a significant effect of time, land cover change and a significant interaction between the two. Dispersion parameter = 0.453

### Check Residuals

```{r model 1.4, check residuals}

# Create tibble for plotting
plot_data <- tibble(
  Fitted = fitted(model1.4_municipality),
  Residuals = residuals(model1.4_municipality))

# Plot the residuals vs fitted values
ggplot(plot_data, aes(x = Fitted, y = Residuals)) +
  geom_point() +
  theme_classic()
```

### Simulate Residuals

```{r model 1.4, simulate residuals}

# Simulate residuals 
#model1.4_municipality_simulation <- simulateResiduals(fittedModel = model1.4_municipality)

# Plot simulated residuals
#plot(model1.4_municipality_simulation)
```

### Test Overdispersion

```{r model 1.4, check overdispersion}
#testDispersion(model1.4_municipality_simulation) 
```

### Test Zero-Inflation

```{r model 1.4, check zero-inflation}
#testZeroInflation(model1.4_municipality_simulation) 
```

# Choose nbinom1 or nbinom2

## Compare AIC

```{r compare AIC}
# SSB models
AIC(model1.1_SSB, model1.3_SSB)

# Municipality models
AIC(model1.2_municipality, model1.4_municipality)
```

```{r}
par(mfrow = c(1, 2))
plot(residuals(model1.1_SSB), main = "Residuals for nbinom1")
plot(residuals(model1.3_SSB), main = "Residuals for nbinom2")
```

# Model 1.5

## Model Validation

### Convergence Check

```{r model 1.5, convergence check}

# Load model from file
load(here::here("data", "models", "model1.5_SSB.RData"))

# Get summary of the model
summary(model1.5_SSB)
```

Significant effect of time, land cover change, occurrences before land cover change. No significant interaction between cover change, time period 2006-2012 and occurrences before.

### Check Residuals

```{r model 1.5, check residuals}

# Create tibble for plotting
plot_data <- tibble(
  Fitted = fitted(model1.5_SSB),
  Residuals = residuals(model1.5_SSB))

# Plot the residuals vs fitted values
ggplot(plot_data, aes(x = Fitted, y = Residuals)) +
  geom_point() +
  theme_classic()

```

# Model 1.6

## Model Validation

### Convergence Check

```{r model 1.6, convergence check}

# Load model from file
load(here::here("data", "models", "model1.6_municipality.RData"))

# Get summary of the model
summary(model1.6_municipality)
```

Significant effect of time, land cover change, occurrences before land cover change. No significant interaction between cover change, time period 2006-2012 and occurrences before.

### Check Residuals

```{r model 1.6 check residuals}

# Create tibble for plotting
plot_data <- tibble(
  Fitted = fitted(model1.6_municipality),
  Residuals = residuals(model1.6_municipality))

# Plot the residuals vs fitted values
ggplot(plot_data, aes(x = Fitted, y = Residuals)) +
  geom_point() +
  theme_classic()

```
