---
title: "Land Cover Change (Y/N) and Occurrence Records & Time"
author: "Beatrice Trascau"
date: "2025-02-25"
output: html_document
---

```{r}
library(here)
library(dplyr)
library(DHARMa)
library(ggplot2)
library(knitr)
library(bbmle)
```

# Introduction

This markdown outlines the process of modelling the number of occurrence records after a land cover change in relation to land cover changes (Y/N), time and SSB ID, without and with an offset.

The models ran as part of this are as follows:

| Model Name                            | Interaction?                | Marginal Value |
|--------------------------------|------------------------|-----------------|
| YN_model1_SSB_interaction             | cover_change \* time_period | 0.001          |
| YN_model2_SSB_no_interaction          | cover_change + time_period  | 0.001          |
| YN_model3_SSB_interaction_0.1_offset  | cover_change \* time_period | 0.1            |
| YN_model4_SSB_interaction_0.01_offset | cover_change \* time_period | 0.01           |

# YN_model1_SSB_interaction

## Get output

```{r YN model 1}

# Load model
load(here::here("data", "models", "YN_model1_SSB_interaction.RData"))

# Get model summary
YN_model1_summary <- summary(YN_model1_SSB_interaction)

# Convert summary coefficients to df
YN_model_coeff_df <- as.data.frame(YN_model1_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
YN_model_coeff_df[] <- lapply(YN_model_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(YN_model_coeff_df)
```

# YN_model1_SSB_no_interaction

## Get output

```{r YN model 2}

# Load model
load(here::here("data", "models", "YN_model2_SSB_no_interaction.RData"))

# Get model summary
YN_model2_summary <- summary(YN_model2_SSB_no_interaction)

# Convert summary coefficients to df
YN_model2_coeff_df <- as.data.frame(YN_model2_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
YN_model2_coeff_df[] <- lapply(YN_model2_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(YN_model2_coeff_df)
```

## Compare Model 1 and Model 2

```{r compare AICs }
AICctab(YN_model1_SSB_interaction, YN_model2_SSB_no_interaction, base = TRUE)

#                                AICc           dAICc    df
#YN_model1_SSB_interaction    899375941.7         0.0    8 
#YN_model2_SSB_no_interaction 899411298.0     35356.3    6 
```

# Model 3: 0.1 offset

```{r YN model 3}

# Load model
load(here::here("data", "models", "YN_model3_SSB_interaction_0.1_offset.RData"))

# Get model summary
YN_model3_summary <- summary(YN_model3_SSB_interaction_0.1_offset)

# Convert summary coefficients to df
YN_model3_coeff_df <- as.data.frame(YN_model3_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
YN_model3_coeff_df[] <- lapply(YN_model3_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(YN_model3_coeff_df)
```

# Model 4: 0.01 offset

```{r YN model 4}

# Load model
load(here::here("data", "models", "YN_model4_SSB_interaction_0.01_offset.RData"))

# Get model summary
YN_model4_summary <- summary(YN_model4_SSB_interaction_0.01_offset)

# Convert summary coefficients to df
YN_model4_coeff_df <- as.data.frame(YN_model4_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
YN_model4_coeff_df[] <- lapply(YN_model4_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(YN_model4_coeff_df)
```

# Model 5: 0.1 offset and no interaction

```{r YN model 5}

# Load model
load(here::here("data", "models", "YN_model5_SSB_no_interaction_0.1_offset.RData"))

# Get model summary
YN_model5_summary <- summary(YN_model5_SSB_no_interaction_0.1_offset)

# Convert summary coefficients to df
YN_model5_coeff_df <- as.data.frame(YN_model5_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
YN_model5_coeff_df[] <- lapply(YN_model5_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(YN_model5_coeff_df)
```

# AIC Comparison Model 3 and Model 5

```{r model 3 & model 5 AIC}
AICctab(YN_model3_SSB_interaction_0.1_offset, YN_model5_SSB_no_interaction_0.1_offset, base = TRUE)
 
#                                         AICc        dAICc       df
# YN_model3_SSB_interaction_0.1_offset    896897400.1         0.0 8 
# YN_model5_SSB_no_interaction_0.1_offset 896901072.8      3672.7 6 
```
