---
title: "Land Cover Change (Intensification/Extensification) and Occurrence Records & Time"
author: "Beatrice Trascau"
date: "2025-02-25"
output: html_document
---

# **Introduction**

This markdown outlines the process of modelling the number of species occurrence records in relation to land cover changes, time and SSB ID. Here, land cover changes are categorised as representing either a form of intensification or extensification of land-use.

```{r library load, include = FALSE}
library(here)
library(dplyr)
library(DHARMa)
library(ggplot2)
library(bbmle)
```

| Model Name                             | Interaction?                | Marginal Value |
|--------------------------------|-----------------------|-----------------|
| IntensExtens_model1_SSB_interaction    | cover_change \* time_period | 0.001          |
| IntensExtens_model2_SSB_no_interaction | cover_change + time_period  | 0.001          |
| IntensExtens_model3_SSB_0.1_offset     | cover_change \* time_period | 0.1            |
| IntensExtens_model4_SSB_0.01_offset    | cover_change \* time_period | 0.01           |

# IntensExtens_model1_SSB_interaction

## Get output

```{r  IntensExtens model 1}

# Load model
load(here::here("data", "models", "IntensExtense_model1_SSB_interaction.RData"))

# Get model summary
IntensExtens_model1_summary <- summary(IntensExtense_model1_SSB_interaction)

# Convert summary coefficients to df
IntensExtens_model_coeff_df <- as.data.frame(IntensExtens_model1_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
IntensExtens_model_coeff_df[] <- lapply(IntensExtens_model_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(IntensExtens_model_coeff_df)
```

# IntensExtens_model2_SSB_no_interaction

## Get output

```{r IntensExtens model 2}

# Load model
load(here::here("data", "models", "IntensExtens_model2_SSB_no_interaction.RData"))

# Get model summary
IntensExtens_model2_summary <- summary(IntensExtens_model2_SSB_no_interaction)

# Convert summary coefficients to df
IntensExtens_model2_coeff_df <- as.data.frame(IntensExtens_model2_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
IntensExtens_model2_coeff_df[] <- lapply(IntensExtens_model2_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(IntensExtens_model2_coeff_df)
```

## Compare Model 1 and Model 2

```{r compare AICs }
AICctab(IntensExtens_model_coeff_df, IntensExtens_model2_coeff_df, base = TRUE)

#                                AICc           dAICc    df
#YN_model1_SSB_interaction    899375941.7         0.0    8 
#YN_model2_SSB_no_interaction 899411298.0     35356.3    6 
```

# IntensExtens_model3_SSB_0.1_offset

```{r IntensExtens model 3}

# Load model
load(here::here("data", "models", "IntensExtens_model3_SSB_0.1_offset.RData"))

# Get model summary
IntensExtens_model3_summary <- summary(IntensExtens_model3_SSB_0.1_offset)

# Convert summary coefficients to df
IntensExtens_model3_coeff_df <- as.data.frame(IntensExtens_model3_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
IntensExtens_model3_coeff_df[] <- lapply(IntensExtens_model3_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(IntensExtens_model3_coeff_df)
```

# IntensExtens_model4_SSB_0.01_offset

```{r IntensExtens model4}

# Load model
load(here::here("data", "models", "IntensExtens_model4_SSB_0.01_offset.RData"))

# Get model summary
IntensExtens_model4_summary <- summary(IntensExtens_model4_SSB_0.01_offset)

# Convert summary coefficients to df
IntensExtens_model4_coeff_df <- as.data.frame(IntensExtens_model4_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
IntensExtens_model4_coeff_df[] <- lapply(IntensExtens_model4_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(IntensExtens_model4_coeff_df)
```

# IntensExtens_model5_SSB_no_interaction_0.1_offset

```{r IntensExtens model 5}

# Load model
load(here::here("data", "models", "IntensExtens_model5_SSB_no_interaction_0.1_offset.RData"))

# Get model summary
IntensExtens_model5_summary <- summary(IntensExtens_model5_SSB_no_interaction_0.1_offset)

# Convert summary coefficients to df
IntensExtens_model5_coeff_df <- as.data.frame(IntensExtens_model5_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
IntensExtens_model5_coeff_df[] <- lapply(IntensExtens_model5_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(IntensExtens_model5_coeff_df)
```

# Model 3 & Model 5 AIC Comparison

```{r model 3 & model 5 AIC}
AICctab(IntensExtens_model3_SSB_0.1_offset, IntensExtens_model5_SSB_no_interaction_0.1_offset, base = TRUE)
```
