---
title: "Land Cover Change (Types) and Occurrence Records & Time"
author: "Beatrice Trascau"
date: "2024-06-12"
output: html_document
---

# Introduction

This markdown outlines the process of modelling the number of species occurrence records in relation to land cover changes, time and SSB ID (or Municipality). Here, we use the individual types of land cover changes. These models were ran on a 10% subset of the data to save computational power.

The defined model was:

-   occurrences_after \~ cover_change(types) \* time_period + (1\|SSBID) (model3.1 & model3.3)

-   occurrences_after \~ cover_change(types) \* time_period + (1\|municipality) (model3.2 & model3.4)

-   occurrences_after \~ cover_change(types) \* time_period \* occurrences_before + offset(occurrences_before) + (1\|SSBID) (model3.5)

-   occurrences_after \~ cover_change(types) \* time_period \* occurrences_before + offset(occurrences_before) + (1\|municipality) (model3.6)

The models variants ran as part of this are as follows:

| Model Number | Function | Family  | Link | SSBID/Municipality | Offset? |
|--------------|----------|---------|------|--------------------|---------|
| 3.1.         | glmmTMB  | nbinom1 | log  | SSBID              | N       |
| 3.2.         | glmmTMB  | nbinom1 | log  | Municipality       | N       |
| 3.3.         | glmmTMB  | nbinom2 | log  | SSBID              | N       |
| 3.4.         | glmmTMB  | nbinom2 | log  | Municipality       | N       |
| 3.5.         | glmmTMB  | nbinom2 | log  | SSBID              | Y       |
| 3.6.         | glmmTMB  | nbinom2 | log  | Municipality       | Y       |

```{r library load, include = FALSE}
library(here)
library(dplyr)
library(DHARMa)
library(ggplot2)
```

# Data Exploration

## Inspect Data

```{r inspect data}

# Read data in 
load(here::here("data", "derived_data",
                "occ_cover_change_types_before_after_for_model.rda"))

# Check data structure
glimpse(occ_cover_change_types_before_after_for_model)
```

## Check for Outliers

```{r check outliers}

# Barplot of number of occurrences
ggplot(occ_cover_change_types_before_after_for_model , 
       aes(x = time_period, y = ocurrences_after, fill = time_period))+
  geom_boxplot()+
  theme_classic()
```

Relationships between time period and the number of occurrences is hard to interpret when broken down by the type of land cover change. However, when looking at the relationship without considering land cover change types, it does not appear entirely linear, with the number of records peaking in the 2006-2012 period and then decreasing again.

## Check Normality

```{r check normality}

# Plot histogram of number of records by time period
occ_cover_change_types_before_after_for_model |>
  filter(ocurrences_after < 100) |>
  ggplot(aes(x = ocurrences_after)) +
  geom_histogram(binwidth = 2, position = "dodge", 
                 alpha = 0.4) +
  facet_grid(time_period ~ ., scales = "free") +
  theme_classic()
```

## Check Relationships

```{r check relationships}

# Plot relationships
occ_cover_change_types_before_after_for_model |>
  ggplot(aes(x = time_period, y = ocurrences_after)) +
  geom_point() +
  facet_wrap(~cover_change) +
  theme_classic()
```

# Model 3.1.

```{r model 3.1. convergence check}
# Load model from file
load(here::here("data", "models", "model3.1_SSB.RData"))

# Get summary of the model
summary(model3.1_SSB)
```

No significant effect of time and no significant effect of most cover changes. Significant effect of:

-   Sparse vegetation -\> Sparse vegetation = No change

-   Woodland Shrub -\> Woodland Shrub in 2006 - 2012 = No Change in 2006-2012

-   Sparse vegetation -\> Sparse vegetation in 2012 - 2018 = No Change in 2012 -2018

Statistics could not be computed for some of the categories is NA. Presumably because the model had to drop some of the categories in order to be able to compute.

## Check Residuals

```{r model 3.1. check residuals}

# Create tibble for plotting
plot_data <- tibble(
  Fitted = fitted(model3.1_SSB),
  Residuals = residuals(model3.1_SSB))

# Plot the residuals vs fitted values
ggplot(plot_data, aes(x = Fitted, y = Residuals)) +
  geom_point() +
  theme_classic()
```

# Model 3.2.

```{r model 3.2. convergence check}
# Load model from file
load(here::here("data", "models", "model3.2_municipality.RData"))

# Get summary of the model
summary(model3.2_municipality)
```

No significant effect of time and no significant effect of most cover changes. Significant effect of:

-   Forest -\> Forest = No change

-   Moors, Heathland & Grassland -\> Moors, Heathland & Grassland = No Change

-   Sparse vegetation -\> Sparse vegetation = No change

-   Woodland Shrub -\> Woodland Shrub in 2006 - 2012 = No Change in 2006-2012

-   Sparse vegetation -\> Sparse vegetation in 2012 - 2018 = No Change in 2012 -2018

Statistics could not be computed for some of the categories is NA. Presumably because the model had to drop some of the categories in order to be able to compute.

## Check Residuals

```{r model 3.2. check residuals}

# Create tibble for plotting
plot_data <- tibble(
  Fitted = fitted(model3.2_municipality),
  Residuals = residuals(model3.2_municipality))

# Plot the residuals vs fitted values
ggplot(plot_data, aes(x = Fitted, y = Residuals)) +
  geom_point() +
  theme_classic()
```

# Model 5.3.

```{r model 3.3. convergence check}
# Load model from file
load(here::here("data", "models", "model3.3_SSB.RData"))

# Get summary of the model
summary(model3.3_SSB)
```

Significant effects of:

-   Complex agriculture -\> complex agriculture = No change

-   Complex agriculture -\> urban areas

-   Forests -\> complex agriculture

-   Forests -\> Woodland Shrub

-   Moors, Heathland & Grassland -\> Moors, Heathland & Grassland = No Change

-   Moors, Heathland & Grassland -\> Urban Areas

-   Other -\> Forests

-   Other -\> Other

-   Sparse Vegetation -\> Sparse Vegetation = No Change

-   Sparse Vegetation -\> Urban

-   Urban -\> Urban = No Change

-   Woodland Shrub -\> Urban

-   Woodland Shrub -\> Woodland Shrub = No Change

-   Significant effect of time

-   Complex agriculture -\> Complex Agriculture in 2006-2012

-   Forests -\> Forests in 2006-2012 and 2012-2018

-   Forests -\> Woodland Shrub in 2006-2012

-   Heath, Moors & Grassland -\> Heath, Moors & Grassland in 2006-2012 and 2012-2018

-   Sparse Vegetation -\> Sparse Vegetation in 2006-2012 and 2012-2018

-   Urban -\> Urban in 2006-2012 and 2012-2018

-   Woodland Shrub -\> Forests in 2006-2012 and 2012-2018

-   Woodland Shrub -\> Woodland Shrub in 2006-2012 and 2012-2018

## Check Residuals

```{r model 3.3. check residuals}

# Create tibble for plotting
plot_data <- tibble(
  Fitted = fitted(model3.3_SSB),
  Residuals = residuals(model3.3_SSB))

# Plot the residuals vs fitted values
ggplot(plot_data, aes(x = Fitted, y = Residuals)) +
  geom_point() +
  theme_classic()
```

# Model 3.4.

```{r model 3.4. convergence check}
# Load model from file
load(here::here("data", "models", "model3.4_municipality.RData"))

# Get summary of the model
summary(model3.4_municipality)
```

Significant effects of:

-   Complex agriculture -\> complex agriculture = No change

-   Complex agriculture -\> urban areas

-   Forests -\> complex agriculture

-   Foests -\> Forests

-   Forests -\> Woodland Shrub

-   Other -\> Forests

-   Other -\> Other

-   Sparse Vegetation -\> Sparse Vegetation = No Change

-   Sparse Vegetation -\> Urban

-   Urban -\> Urban = No Change

-   Woodland Shrub -\> Forsts

-   Woodland Shrub -\> Urban

-   Woodland Shrub -\> Woodland Shrub = No Change

-   Significant effect of time

-   Complex agriculture -\> Complex Agriculture in 2006-2012 and 2012-2018

-   Forests -\> Forests in 2006-2012 and 2012-2018

-   Forests -\> Woodland Shrub in 2006-2012

-   Urban -\> Urban in 2006-2012 and 2012-2018

-   Woodland Shrub -\> Forests in 2006-2012 and 2012-2018

-   Woodland Shrub -\> Urban in 2006-2012

-   Woodland Shrub -\> Woodland Shrub in 2006-2012 and 2012-2018

-   Moors, Heath & Grassland -\> Moors, Heath & Grassland in 2012-2018

## Check Residuals

```{r model 3.4. check residuals}

# Create tibble for plotting
plot_data <- tibble(
  Fitted = fitted(model3.4_municipality),
  Residuals = residuals(model3.4_municipality))

# Plot the residuals vs fitted values
ggplot(plot_data, aes(x = Fitted, y = Residuals)) +
  geom_point() +
  theme_classic()
```

# Model 3.5.

```{r model 3.5. convergence check}
# Load model from file
load(here::here("data", "models", "model3.5_SSB.RData"))

# Get summary of the model
summary(model3.5_SSB)
```

Significant effects of:

-   Complex agriculture -\> complex agriculture = No change (also in 2006-2012 and 2012-2018, occurrences before, 2006-2012 & occurrences before, 2012-2018 & occurrences before)

-   Forests -\> forests = No Change (also in 2012-2018, occurrences before, 2006-2012 & occurrences before)

-   Sparse vegetation -\> Sparse Vegetation = No Change (also in 2012-2018, occurrences before, 2006-2012 & occurrences before)

-   Urban -\> Urban = No Change (also in 2006-2012 and 2012-2018, occurrences before, 2006-2012 & occurrences before)

-   Woodland Shrub -\> Woodland Shrub = No Change (also in 2012-2018, occurrences before, 2006-2012 & occurrences before, 2012-2018 & occurrences before)

-   Time period 2006-2012

-   Occurrences before

-   Time period 2006-2012 and Occurrences before

-   Time period 2012-2018 and Occurrences before

-   Moors, Heathland and Grassland -\> Moors, Heathland and Grassland in 2006-2012 & occurrences before, 2012-2018 & occurrences before

## Check Residuals

```{r model 3.5. check residuals}

# Create tibble for plotting
plot_data <- tibble(
  Fitted = fitted(model3.5_SSB),
  Residuals = residuals(model3.5_SSB))

# Plot the residuals vs fitted values
ggplot(plot_data, aes(x = Fitted, y = Residuals)) +
  geom_point() +
  theme_classic()
```

Why are there only 2 data points?

# Model 3.6.

```{r model 3.6. convergence check}
# Load model from file
load(here::here("data", "models", "model3.6_municipality.RData"))

# Get summary of the model
summary(model3.6_municipality)
```

Significant effect of:

-   Complex agriculture -\> Complex agriculture

    -   Including occurrences before

    -   In 2006-2012 & occurrences before

-   Forests -\> Forests

    -   In 2006-2012

    -   In 2012-2018

-   Moors, Heath & Grassland -\> Moors, Heath & Grassland

    -   In 2012-2018

    -   Including occurrences before

    -   In 2006-2012 & occurrences before

    -   In 2012-2018 & occurrences before

-   Sparse vegetation -\> Sparse vegetation

    -   In 2012-2018

    -   In 2012-2018 & occurrences before

-   Urban -\> Urban

    -   In 2006-2012

    -   In 2012-2018

    -   Including occurrences before

    -   In 2006-2012 & occurrences before

    -   In 2012-2018 & occurrences before

-   Woodland Shrub -\> Woodland Shrub

    -   Including occurrences before

    -   In 2006-2012 & occurrences before

    -   In 2012-2018 & occurrences before

-   Time period 2006-2012

    -   Including occurrences before

-   Time period 2012-2018

    -   Including occurrences before

-   Occurrences before

## Check Residuals

```{r model 3.6. check residuals}

# Create tibble for plotting
plot_data <- tibble(
  Fitted = fitted(model3.6_municipality),
  Residuals = residuals(model3.6_municipality))

# Plot the residuals vs fitted values
ggplot(plot_data, aes(x = Fitted, y = Residuals)) +
  geom_point() +
  theme_classic()
```

Again there only seem to be two data points. This needs further investigation.
