---
title: "4.4_cover_change_types_occ_individual_model_exploration"
output: html_document
date: "2025-02-25"
---

# Introduction

This markdown outlines the process of modelling the number of species occurrence records in relation to land cover changes, time and SSB ID (or Municipality). Here, land cover changes are modelled as individual land cover types and compared to a baseline of no change (in that specific land cover type).

```{r library load, include = FALSE}
library(here)
library(dplyr)
library(DHARMa)
library(ggplot2)
library(bbmle)
library(kableExtra)
```

# 1. Urban Fabric

## Model 1: Interaction

```{r  Urban Fabric model 1}

# Load model
load(here::here("data", "models", "urban_model1_SSB_interaction.RData"))

# Get model summary
urban_model1_summary <- summary(urban_model1_SSB_interaction)

# Convert summary coefficients to df
urban_model1_coeff_df <- as.data.frame(urban_model1_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
urban_model1_coeff_df[] <- lapply(urban_model1_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(urban_model1_coeff_df)
```

## Model 2: No Interaction

```{r  Urban Fabric model 2}

# Load model
load(here::here("data", "models", "urban_model2_SSB_no_interaction.RData"))

# Get model summary
urban_model2_summary <- summary(urban_model2_SSB_no_interaction)

# Convert summary coefficients to df
urban_model2_coeff_df <- as.data.frame(urban_model2_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
urban_model2_coeff_df[] <- lapply(urban_model2_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(urban_model2_coeff_df)
```

## Compare Model 1 and Model 2

```{r urban compare AICs }
AICctab(urban_model1_SSB_interaction, urban_model2_SSB_no_interaction, base = TRUE) 

```

## Model 3: 0.1 marginal value

```{r urban fabric model 3}

# Load model
load(here::here("data", "models", "urban_model3_SSB_interaction_0.1_offset.RData"))

# Get model summary
urban_model3_summary <- summary(urban_model3_SSB_interaction_0.1_offset)

# Convert summary coefficients to df
urban_model3_coeff_df <- as.data.frame(urban_model3_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
urban_model3_coeff_df[] <- lapply(urban_model3_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(urban_model3_coeff_df)
```

## Model 4: 0.01 marginal value

```{r urban fabric model 4}

# Load model
load(here::here("data", "models", "urban_model4_SSB_interaction_0.01_offset.RData"))

# Get model summary
urban_model4_summary <- summary(urban_model4_SSB_interaction_0.01_offset)

# Convert summary coefficients to df
urban_model4_coeff_df <- as.data.frame(urban_model4_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
urban_model4_coeff_df[] <- lapply(urban_model4_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(urban_model4_coeff_df)
```

## Model 5: 0.1 marginal value and no interaction

```{r urban model 5}

# Load model
load(here::here("data", "models", "urban_model5_SSB_no_interaction_0.1_offset.RData"))

# Get model summary
urban_model5_summary <- summary(urban_model5_SSB_no_interaction_0.1_offset)

# Convert summary coefficients to df
urban_model5_coeff_df <- as.data.frame(urban_model5_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
urban_model5_coeff_df[] <- lapply(urban_model5_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(urban_model5_coeff_df)
```

## Compare AIC Model 3 & Model 5

```{r urban model 3 & 5 AIC}

AICctab(urban_model3_SSB_interaction_0.1_offset, urban_model5_SSB_no_interaction_0.1_offset, base = TRUE)

#                                            AICc       dAICc      df
# urban_model5_SSB_no_interaction_0.1_offset 87844894.6        0.0 7 
# urban_model3_SSB_interaction_0.1_offset    87844895.0        0.4 9 
```

# Complex Agricultural Cover

## Model 1: Interaction

```{r CAC model 1}

# Load model
load(here::here("data", "models", "complex_agri_model1_SSB_interaction.RData"))

# Get model summary
ca_model1_summary <- summary(complex_agri_model1_SSB_interaction)

# Convert summary coefficients to df
ca_model1_coeff_df <- as.data.frame(ca_model1_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
ca_model1_coeff_df[] <- lapply(ca_model1_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(ca_model1_coeff_df)
```

## Model 2: No Interaction

```{r CAC model 2}

# Load model
load(here::here("data", "models", "complex_agri_model2_SSB_no_interaction.RData"))

# Get model summary
ca_model2_summary <- summary(complex_agri_model2_SSB_no_interaction)

# Convert summary coefficients to df
ca_model2_coeff_df <- as.data.frame(ca_model2_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
ca_model2_coeff_df[] <- lapply(ca_model2_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(ca_model2_coeff_df)

```

## Compare Model 1 and Model 2

```{r compare AICs }
AICctab(complex_agri_model1_SSB_interaction, complex_agri_model2_SSB_no_interaction, base = TRUE)  
```

## Model 3: 0.1 marginal value

```{r CAC model 3}
# Load model
load(here::here("data", "models", "complex_agri_model3_SSB_interaction_0.1_offset.RData"))

# Get model summary
ca_model3_summary <- summary(complex_agri_model3_SSB_interaction_0.1_offset)

# Convert summary coefficients to df
ca_model3_coeff_df <- as.data.frame(ca_model3_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
ca_model3_coeff_df[] <- lapply(ca_model3_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(ca_model3_coeff_df)

```

## Model 4: 0.01 marginal value

```{r CAC model 3}

# Load model
load(here::here("data", "models", "complex_agri_model4_SSB_interaction_0.01_offset.RData"))

# Get model summary
ca_model4_summary <- summary(complex_agri_model4_SSB_interaction_0.01_offset)

# Convert summary coefficients to df
ca_model4_coeff_df <- as.data.frame(ca_model4_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
ca_model4_coeff_df[] <- lapply(ca_model4_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(ca_model4_coeff_df)

```

## Model 5: 0.1 & no interaction

```{r CAC model 5}

# Load model
load(here::here("data", "models", "complex_agri_model5_SSB_no_interaction_0.1_offset.RData"))

# Get model summary
ca_model5_summary <- summary(complex_agri_model5_SSB_no_interaction_0.1_offset)

# Convert summary coefficients to df
ca_model5_coeff_df <- as.data.frame(ca_model5_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
ca_model5_coeff_df[] <- lapply(ca_model5_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(ca_model5_coeff_df)
```

## Compare AIC Model 3 & Model 5

```{r CAC model 3 & 5 compare AIC}
AICctab(complex_agri_model3_SSB_interaction_0.1_offset, complex_agri_model5_SSB_no_interaction_0.1_offset, base = TRUE) 

#                                                   AICc       dAICc      df
# complex_agri_model3_SSB_interaction_0.1_offset    23296033.4        0.0 16
# complex_agri_model5_SSB_no_interaction_0.1_offset 23296119.2       85.8 10
```

# Agriculture with Significant Natural Vegetation

## Model 1: Interaction

```{r ASNV model 1}

# Load model
load(here::here("data", "models", "agri_veg_model1_SSB_interaction.RData"))

# Get model summary
asnv_model1_summary <- summary(agri_veg_model1_SSB_interaction)

# Convert summary coefficients to df
asnv_model1_coeff_df <- as.data.frame(asnv_model1_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
asnv_model1_coeff_df[] <- lapply(asnv_model1_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(asnv_model1_coeff_df)
```

## Model 2: No Interaction

```{r ASNV model 2}

# Load model
load(here::here("data", "models", "agri_veg_model2_SSB_no_interaction.RData"))

# Get model summary
asnv_model2_summary <- summary(agri_veg_model2_SSB_no_interaction)

# Convert summary coefficients to df
asnv_model2_coeff_df <- as.data.frame(asnv_model2_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
asnv_model2_coeff_df[] <- lapply(asnv_model2_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(asnv_model2_coeff_df)
```

## Compare Model 1 and Model 2

```{r compare AICs }

AICctab(agri_veg_model1_SSB_interaction, agri_veg_model2_SSB_no_interaction, base = TRUE)  
```

## Model 3: 0.1 marginal value

```{r ASNV model 3}

# Load model
load(here::here("data", "models", "agri_veg_model3_SSB_interaction_0.1_offset.RData"))

# Get model summary
asnv_model3_summary <- summary(agri_veg_model3_SSB_interaction_0.1_offset)

# Convert summary coefficients to df
asnv_model3_coeff_df <- as.data.frame(asnv_model3_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
asnv_model3_coeff_df[] <- lapply(asnv_model3_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(asnv_model3_coeff_df)
```

## Model 4: 0.01 marginal value

```{r ASNV model 4}

# Load model
load(here::here("data", "models", "agri_veg_model4_SSB_interaction_0.01_offset.RData"))

# Get model summary
asnv_model4_summary <- summary(agri_veg_model4_SSB_interaction_0.01_offset)

# Convert summary coefficients to df
asnv_model4_coeff_df <- as.data.frame(asnv_model4_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
asnv_model4_coeff_df[] <- lapply(asnv_model4_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(asnv_model4_coeff_df)

```

## Model 5: 0.1 & no interaction

```{r ASNV model 5}

# Load model
load(here::here("data", "models", "agri_veg_model5_SSB_no_interaction_0.1_offset.RData"))

# Get model summary
asnv_model5_summary <- summary(agri_veg_model5_SSB_no_interaction_0.1_offset)

# Convert summary coefficients to df
asnv_model5_coeff_df <- as.data.frame(asnv_model5_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
asnv_model5_coeff_df[] <- lapply(asnv_model5_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(asnv_model5_coeff_df)
```

## Compare AIC Model 3 & Model 5

```{r ASNV model 3 & 5 AIC}
AICctab(agri_veg_model3_SSB_interaction_0.1_offset, agri_veg_model5_SSB_no_interaction_0.1_offset, base = TRUE)

#                                               AICc       dAICc      df
# agri_veg_model3_SSB_interaction_0.1_offset    18256470.5        0.0 17
# agri_veg_model5_SSB_no_interaction_0.1_offset 18256490.6       20.1 10
```

# Forests

## Model 1: Interaction

```{r forests model 1}

# Load model
load(here::here("data", "models", "forests_model1_SSB_interaction.RData"))

# Get model summary
forests_model1_summary <- summary(forests_model1_SSB_interaction)

# Convert summary coefficients to df
forests_model1_coeff_df <- as.data.frame(forests_model1_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
forests_model1_coeff_df[] <- lapply(forests_model1_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(forests_model1_coeff_df)
```

## Model 2: No interaction

```{r forests model 2}

# Load model
load(here::here("data", "models", "forests_model2_SSB_no_interaction.RData"))

# Get model summary
forests_model2_summary <- summary(forests_model2_SSB_no_interaction)

# Convert summary coefficients to df
forests_model2_coeff_df <- as.data.frame(forests_model2_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
forests_model2_coeff_df[] <- lapply(forests_model2_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(forests_model2_coeff_df)
```

## Compare Model 1 and Model 2

```{r forests compare AIC}
AICctab(forests_model1_SSB_interaction, forests_model2_SSB_no_interaction, base = TRUE)  
```

## Model 3: 0.1 marginal value

```{r forests model 3}

# Load model
load(here::here("data", "models", "forests_model3_SSB_interaction_0.1_offset.RData"))

# Get model summary
forests_model3_summary <- summary(forests_model3_SSB_interaction_0.1_offset)

# Convert summary coefficients to df
forests_model3_coeff_df <- as.data.frame(forests_model3_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
forests_model3_coeff_df[] <- lapply(forests_model3_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(forests_model3_coeff_df)
```

## Model 4: 0.01 marginal value

```{r forests model 4}

# Load model
load(here::here("data", "models", "forests_model4_SSB_interaction_0.01_offset.RData"))

# Get model summary
forests_model4_summary <- summary(forests_model4_SSB_interaction_0.01_offset)

# Convert summary coefficients to df
forests_model4_coeff_df <- as.data.frame(forests_model4_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
forests_model4_coeff_df[] <- lapply(forests_model4_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(forests_model4_coeff_df)
```

## Model 5: 0.1 & no interaction

```{r forests model 5}

# Load model
load(here::here("data", "models", "forests_model5_SSB_no_interaction_0.1_offset.RData"))

# Get model summary
forests_model5_summary <- summary(forests_model5_SSB_no_interaction_0.1_offset)

# Convert summary coefficients to df
forests_model5_coeff_df <- as.data.frame(forests_model5_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
forests_model5_coeff_df[] <- lapply(forests_model5_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(forests_model5_coeff_df)
```

## Compare AIC Model 3 & Model 5

```{r forests model 3 & 5 AIC}
AICctab(forests_model3_SSB_interaction_0.1_offset, forests_model5_SSB_no_interaction_0.1_offset, base = TRUE)

#                                              AICc        dAICc       df
# forests_model3_SSB_interaction_0.1_offset    201983553.0         0.0 23
# forests_model5_SSB_no_interaction_0.1_offset 201984905.5      1352.5 11
```

# Moors, Heathland & Grassland

## Model 1: Interaction

```{r mhg model 1}

# Load model
load(here::here("data", "models", "moors_model1_SSB_interaction.RData"))

# Get model summary
moors_model1_summary <- summary(moors_model1_SSB_interaction)

# Convert summary coefficients to df
moors_model1_coeff_df <- as.data.frame(moors_model1_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
moors_model1_coeff_df[] <- lapply(moors_model1_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(moors_model1_coeff_df)
```

## Model 2: No Interaction

```{r mhg model 2}

# Load model
load(here::here("data", "models", "moors_model2_SSB_no_interaction.RData"))

# Get model summary
moors_model2_summary <- summary(moors_model2_SSB_no_interaction)

# Convert summary coefficients to df
moors_model2_coeff_df <- as.data.frame(moors_model2_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
moors_model2_coeff_df[] <- lapply(moors_model2_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(moors_model2_coeff_df)
```

## Compare Model 1 and Model 2

```{r moors compare models}
AICctab(moors_model1_SSB_interaction, moors_model2_SSB_no_interaction, base = TRUE)
```

## Model 3: 0.1 marginal value

```{r mhg model 3}

# Load model
load(here::here("data", "models", "moors_model3_SSB_interaction_0.1_offset.RData"))

# Get model summary
moors_model3_summary <- summary(moors_model3_SSB_interaction_0.1_offset)

# Convert summary coefficients to df
moors_model3_coeff_df <- as.data.frame(moors_model3_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
moors_model3_coeff_df[] <- lapply(moors_model3_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(moors_model3_coeff_df)
```

## Model 4: 0.01 marginal value

```{r mhg model 4}

# Load model
load(here::here("data", "models", "moors_model4_SSB_interaction_0.01_offset.RData"))

# Get model summary
moors_model4_summary <- summary(moors_model4_SSB_interaction_0.01_offset)

# Convert summary coefficients to df
moors_model4_coeff_df <- as.data.frame(moors_model4_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
moors_model4_coeff_df[] <- lapply(moors_model4_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(moors_model4_coeff_df)
```

## Model 5: 0.1 & no interaction

```{r mhg model 5}

# Load model
load(here::here("data", "models", "moors_model5_SSB_no_interaction_0.1_offset.RData"))

# Get model summary
moors_model5_summary <- summary(moors_model5_SSB_no_interaction_0.1_offset)

# Convert summary coefficients to df
moors_model5_coeff_df <- as.data.frame(moors_model5_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
moors_model5_coeff_df[] <- lapply(moors_model5_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(moors_model5_coeff_df)
```

## Compare AIC Model 3 & Model 5

```{r mhg model 3 & 5 AIC}
AICctab(moors_model3_SSB_interaction_0.1_offset, moors_model5_SSB_no_interaction_0.1_offset, base = TRUE)

#                                            AICc      dAICc     df
# moors_model3_SSB_interaction_0.1_offset    7978336.2       0.0 14
# moors_model5_SSB_no_interaction_0.1_offset 7978367.5      31.3 10
```

# Transitional Woodland Shrub

## Model 1: Interaction

```{r tws model 1}

# Load model
load(here::here("data", "models", "woodland_model1_SSB_interaction.RData"))

# Get model summary
woodland_model1_summary <- summary(woodland_model1_SSB_interaction)

# Convert summary coefficients to df
woodland_model1_coeff_df <- as.data.frame(woodland_model1_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
woodland_model1_coeff_df[] <- lapply(woodland_model1_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(woodland_model1_coeff_df)
```

## Model 2: No Interaction

```{r tws model 2}

# Load model
load(here::here("data", "models", "woodland_model2_SSB_no_interaction.RData"))

# Get model summary
woodland_model2_summary <- summary(woodland_model2_SSB_no_interaction)

# Convert summary coefficients to df
woodland_model2_coeff_df <- as.data.frame(woodland_model2_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
woodland_model2_coeff_df[] <- lapply(woodland_model2_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(woodland_model2_coeff_df)
```

## Compare Model 1 and Model 2

```{r tws model comparison}
AICctab(woodland_model1_SSB_interaction, woodland_model2_SSB_no_interaction, base = TRUE)
```

## Model 3: 0.1 marginal value

```{r tws model 3}

# Load model
load(here::here("data", "models", "woodland_model3_SSB_interaction_0.1_offset.RData"))

# Get model summary
woodland_model3_summary <- summary(woodland_model3_SSB_interaction_0.1_offset)

# Convert summary coefficients to df
woodland_model3_coeff_df <- as.data.frame(woodland_model3_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
woodland_model3_coeff_df[] <- lapply(woodland_model3_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(woodland_model3_coeff_df)
```

## Model 4: 0.01 marginal value

```{r tws model 4}

# Load model
load(here::here("data", "models", "woodland_model4_SSB_interaction_0.01_offset.RData"))

# Get model summary
woodland_model4_summary <- summary(woodland_model4_SSB_interaction_0.01_offset)

# Convert summary coefficients to df
woodland_model4_coeff_df <- as.data.frame(woodland_model4_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
woodland_model4_coeff_df[] <- lapply(woodland_model4_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(woodland_model4_coeff_df)
```

## Model 5: 0.1 & no interaction

```{r tws model 5}

# Load model
load(here::here("data", "models", "woodland_model5_SSB_no_interaction_0.1_offset.RData"))

# Get model summary
woodland_model5_summary <- summary(woodland_model5_SSB_no_interaction_0.1_offset)

# Convert summary coefficients to df
woodland_model5_coeff_df <- as.data.frame(woodland_model5_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
woodland_model5_coeff_df[] <- lapply(woodland_model5_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(woodland_model5_coeff_df)
```

## Compare AIC Model 3 & Model 5

```{r tws model 3 & 5 AIC}
AICctab(woodland_model3_SSB_interaction_0.1_offset, woodland_model5_SSB_no_interaction_0.1_offset, base = TRUE)

#                                               AICc     dAICc    df
# woodland_model3_SSB_interaction_0.1_offset    968593.4      0.0 14
# woodland_model5_SSB_no_interaction_0.1_offset 969440.7    847.2 10
```

# Sparse Vegetation

## Model 1: Interaction

```{r sv model 1}

# Load model
load(here::here("data", "models", "sparse_model1_SSB_interaction.RData"))

# Get model summary
sparse_model1_summary <- summary(sparse_model1_SSB_interaction)

# Convert summary coefficients to df
sparse_model1_coeff_df <- as.data.frame(sparse_model1_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
sparse_model1_coeff_df[] <- lapply(sparse_model1_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(sparse_model1_coeff_df)
```

## Model 2: No Interaction

```{r sv model 2}

# Load model
load(here::here("data", "models", "sparse_model2_SSB_no_interaction.RData"))

# Get model summary
sparse_model2_summary <- summary(sparse_model2_SSB_no_interaction)

# Convert summary coefficients to df
sparse_model2_coeff_df <- as.data.frame(sparse_model2_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
sparse_model2_coeff_df[] <- lapply(sparse_model2_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(sparse_model2_coeff_df)
```

## Compare Model 1 and Model 2

```{r sv compare models}
AICctab(sparse_model1_SSB_interaction, sparse_model2_SSB_no_interaction, base = TRUE)
```

## Model 3: 0.1 marginal value

```{r sv model 3}

# Load model
load(here::here("data", "models", "sparse_model3_SSB_interaction_0.1_offset.RData"))

# Get model summary
sparse_model3_summary <- summary(sparse_model3_SSB_interaction_0.1_offset)

# Convert summary coefficients to df
sparse_model3_coeff_df <- as.data.frame(sparse_model3_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
sparse_model3_coeff_df[] <- lapply(sparse_model3_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(sparse_model3_coeff_df)
```

## Model 4: 0.01 marginal value

```{r sv model 4}

# Load model
load(here::here("data", "models", "sparse_model4_SSB_interaction_0.01_offset.RData"))

# Get model summary
sparse_model4_summary <- summary(sparse_model4_SSB_interaction_0.01_offset)

# Convert summary coefficients to df
sparse_model4_coeff_df <- as.data.frame(sparse_model4_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
sparse_model4_coeff_df[] <- lapply(sparse_model4_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(sparse_model4_coeff_df)
```

## Model 5: 0.1 & no interaction

```{r sv model 5}

# Load model
load(here::here("data", "models", "sparse_model5_SSB_no_interaction_0.1_offset.RData"))

# Get model summary
sparse_model5_summary <- summary(sparse_model5_SSB_no_interaction_0.1_offset)

# Convert summary coefficients to df
sparse_model5_coeff_df <- as.data.frame(sparse_model5_summary$coefficients$cond)

# Make sure coefficients are using decimal format (not scientific)
sparse_model5_coeff_df[] <- lapply(sparse_model5_coeff_df, 
                              function(x) format(x, scientific = FALSE, digits = 5))

# Print table using knitr::kable
kable(sparse_model5_coeff_df)
```

## Compare AIC Model 3 & Model 5

```{r sv model 3 & 5 AIC}
AICctab(sparse_model3_SSB_interaction_0.1_offset, sparse_model5_SSB_no_interaction_0.1_offset, base = TRUE)

#                                             AICc      dAICc     df
# sparse_model3_SSB_interaction_0.1_offset    5057541.6       0.0 12
# sparse_model5_SSB_no_interaction_0.1_offset 5057546.1       4.5 8 
```

# Differences

| Cover Type                                       | Variable                                             | 0.001         | 0.1           | 0.01          |
|---------------|---------------|---------------|---------------|---------------|
| Agricultural with Significant Natural Vegetation | 2012-2018                                            | negative (S)  | positive (NS) | negatives (S) |
| Agricultural with Significant Natural Vegetation | ASNV -\> Urban \* 2012-2018                          | negative (S)  | negative (S)  | positive (NS) |
| Forests                                          | Forest -\> Moors, Heathland & Grassland              | positive (S)  | negative (NS) | positive (NS) |
| Moors, Heathland & Grassland                     | Moors, Heathland & Grassland -\> Complex Agriculture | positive (S)  | negative (NS) | positive (NS) |
| Moors, Heathland & Grassland                     | Moors, Heathland & Grassland -\> Forest              | positive (NS) | negative (S)  | positive (NS) |
| Transitional Woodland Shrub                      | Transitional Woodland Shrub -\> Urban \* 06-12       | negative (S)  | positive (NS) | negative (NS) |
| Sparse Vegetation                                | Sparse Vegetation -\> Forests                        | negative (NS) | positive (S)  | positive (S)  |
| Sparse Vegetation                                | Sparse Vegetation -\> Urban                          | positive (S)  | negative (NS) | positive (S)  |
